---
title: "ANLY 510 Assignment 2"
author: "Akane Simpson"
date: "11/21/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

  In the Espresso data there are three methods of espresso brewing that we observed.
The methods were split into three groups of 9 samples. The creme levels were measured between the three groups. A research question that may be proposed is this: 'is there a significant difference in espresso's creme levels between the three brewing methods?'


## Hypothesis

  From the research question we can establish a null and alternative hypothesis in regards to the espresso levels. 
  
  - $H_O$: There is no significant difference in espresso creme levels between the brewing methods. 
  - $H_A$: There is at least one brewing method that has a signifcant difference in espresso creme levels.

## R Code

```{r cars, eval=FALSE}
library(moments)

master = read.csv('./EspressoData.csv')

data = master
summary(data)

data$brewmethod = factor(data$brewmethod)

# Skewness
plot(density(data$cereme))
agostino.test(data$cereme)
# The density plot and Agostino test is used for the skewness test.
# The result showed the that the p-value not being less that the significant 
# value of 0.05 (p< 0.05). We can assume that the sample does not have skewness.

# Normality
qqnorm(data$cereme)
shapiro.test(data$cereme)
# The qqnorm plot and Shpairo Test is used for normality test.
# The resuslts shows that the Espresso data is normal and significant. 
# With a p < 0.05 we can reject the assumption that the Espresso data is normal.


# Independence
espresso.lm = lm(cereme ~ brewmethod, data = data)
espresso.res = resid(espresso.lm)

plot(data$cereme, espresso.res, xlab = "Creme Level", ylab = "Residual", 
     main = "Residual against Creme Level")
abline(0,0)
# The residual plot shows that the values are mostly spread across the graph
# evenly.


# Equal Variance
bartlett.test(data$cereme, data$brewmethod)
tapply(data$cereme, data$brewmethod, var)
# The Bartlett test checks for variance in the data. Since the p< 0.05 we can 
# assume that it passes the variance assumption


# ANOVA Analysis
espresso.aov = aov(cereme ~ brewmethod, data = data)
summary(espresso.aov)
# The summary of the ANOVA model shows that the brewing method does have a significant effect
# on the creme levels of the espressos. 


# Post Hoc
library(pgirmess)

TukeyHSD(espresso.aov)
# We use the Tukey HSD test to find statistical difference between the three 
# brewing methods

# Effect Size
library(pastecs)
library(compute.es)

by(data$cereme, data$brewmethod, stat.desc)
# The by() and stat.desc method shows a descriptive summary of values for each 
# brewing method group.

# Method 1 & 2
mes(61.3,32.4,10.1,7.3,9,9)

# Method 2 & 3
mes(61.3,39.7,10.1,7.7,9,9)

# Method 1 & 3
mes(39.7,32.4,7.7,7.3,9,9)

# The mes() method is used to return effect values between two groups. In this
# we run several mes methods between each group of brewing methods.
```

## APA Summary

  Observations from the study were analyzed by conducting a one-way analysis of variance using R version 4.1.1. First, all assumptions are met and there is no adjustment made. Results suggest that the creme level of the espresso was affected by its brewing method (as 1, 2 or 3) of the participants (F(2, 24) = 28.41, p < .001)
	
	Continue the discussion with specifically which groups differed, a Tukey’s hoc test was established. The result suggested that there is a significant difference between brewing method 1 and brewing method 2 (p < .001) and bewing method 1 and 3 (p < .001), in terms of the valuation of creme levels. There was no significant difference between brewing methods 2 and 3. The effect was large between methods 1 & 2, and 1 & 3, Cohen’s D = 3.28 for methods 1 and 2 and Cohen's D = 0.97 for methods 1 and 3. The Cohen's D for methods 2 and 3 was D = 2.41. 

