---
title: 'Logistic Regression'
author: "STUDENT NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Libraries + Functions

Load all the libraries or functions that you will use to for the rest of the assignment. It is helpful to define your libraries and functions at the top of a report, so that others can know what they need for the report to compile correctly.

Understanding whether a written review is positive or negative can be tricky as the context of what is being reviewed and other factors can impact the sentiment of a review. In this assignment, you will investigate if the words used in a reviews can predict their sentiment. The datasets come from a Kaggle project with labelled sentences which you can check out here [https://www.kaggle.com/marklvl/sentiment-labelled-sentences-data-set].

For this assignment, you can pick one of three datasets to analyze: Amazon Reviews, Yelp Reviews, or Movie Reviews. The first column in each dataset is a measure of sentiment of the review (0 = negative, 1 = positive) and the second is the number of tokens (or words) in the review. The rest of the columns are words that were either used (coded as 1) or not used (coded at 0) in the review. The sentiment of the review should be used as your outcome in your binary logistic regression. For your predictors, choose 10-20 words to test if use of those words predicts the sentiment of the review. 

```{r libraries}
#In using VM, you may need to install the packages 'gridExtra', 'htmlwidgets', and 'readxl' first
```

## Choosing Predictors
- To create a simple, easy-to-interpret model, pick 10-20 words from one of the dataset which capture features most important to a good product (Amazon reviews), restaurant (Yelp reviews) or movie (Movie Reviews). Think about what words are strongly associated with either positive or negative reviews as well as words that may have different meanings in a positive versus negative review.
- Which words did you choose as predictors and why?

  - ANSWER: 

## Running a Binary Logistic Regression

- Run the logistic regression using the `rms` package. (Answer all questions below after ANSWER: using complete sentences)
  - Use the $\chi^2$ test - is the overall model predictive of sentiment? Is it significant?
  - What is Nagelkerke's pseudo-$R^2$? What does it tell you about goodness of fit?
  - What is the C statistic? How well are you predicting sentiment?

  - ANSWER: 
  
```{r runrms}

```

## Coefficients

- Explain each coefficient (i.e. each word you chose) - are they significant? What do they imply if they are significant (i.e., which sentiment does it predict)? (Should be at least a paragraph.)

  - ANSWER: 

## Variable Selection

- Use *either* a backwards stepwise approach or the drop 1 approach to determine which predictors you would keep in the model. 
- Fit the model with glm function in order to test either approach (but NOT both)
- Which predictors would you retain?

  - ANSWER: 

```{r varselect}

```

## Outliers
- Use the model with glm function in order to test assumptions
- Use the `car` library and the `influencePlot()` to create a picture of the outliers. 
  - Are there major outliers for this data?
    - ANSWER:

```{r outliers}

```

## Assumptions

- Explore the `vif` values of the model and determine if you meet the assumption of additivity (meaning no multicollinearity). 
- Are there issues with multicollinearity?
  - ANSWER: 

```{r vif}

```

## Test for Overfitting
- Use the validate function to test for overfitting.
- Is there evidence of overfitting?
  - ANSWER: 

```{r overfittest}

```

## Discussion Question

- Describe a set of texts and research question that interests you that could be explored using this method. Basically, what is a potential application of this method to another area of research?
  
  - ANSWER: 