---
title: "Lecture_8_Assignment_8"
author: "Akane Simpson"
date: "9/23/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("atheism.RData")
```

# Inference for categorical data

In August of 2012, news outlets ranging from the Washington Post to the Huffington Post ran a story about the rise of atheism in America. The source for the story was a poll that asked people, "Irrespective of whether you attend a place of worship or not, would you say you are a religious person, not a religious person or a convinced atheist?" 
This type of question, which asks people to classify themselves in one way or another, is common in polling and generates categorical data. In this lab we take a look at the atheism survey and explore what's at play when making inference about population proportions using categorical data.

The question of atheism was asked by WIN-Gallup International in a similar survey that was conducted in 2005. (We assume here that sample sizes have remained the same.) Table 4 on page 13 of the report summarizes survey results from 2005 and 2012 for 39 countries.

1. Answer the following two questions using the inference function. As always, write out the hypotheses for any tests you conduct and outline the status of the conditions for inference.

  a. Is there convincing evidence that Spain has seen a change in its atheism index between 2005 and 2012?

  Hint: Create a new data set for respondents from Spain. Form confidence intervals for the true proportion of athiests in both years, and determine whether they overlap.
```{r}
# Ho: There is no difference in the atheist population in Spain since 2005 to 2012 (Ho: p.1 == p.2 )
# Ha: There is a difference in the atheist population in Spain since 2005 to 2012 (Ha: p.1 != p.2)
spain05 <- subset(atheism, nationality == "Spain" & year == 2005)
spain12 <- subset(atheism, nationality == "Spain" & year == 2012)

# Next, prove that the data set fulfills the proportion law
table(spain05$response)
table(spain12$response)

inference(spain05$response, success = "atheist", type = "ci", method = "theoretical", est = "proportion")

inference(spain12$response, success = "atheist", type = "ci", method = "theoretical", est = "proportion")
```


```{r, echo=FALSE}
print("The confidence interval for Spain in 2005 was between (0.083:0.1177) while the 
confidence interval in 2012 was between (0.0734:0.1065). The two intervals overlap and shows no significant change in the atheist population between the years. Therefore we cannot reject the null hypothesis.")
```

 
  b. Is there convincing evidence that the United States has seen a change in its atheism index between 2005 and 2012?
```{r}
# Ho: There is no difference in the atheist population in US since 2005 to 2012 (Ho: p.1 == p.2 )
# Ha: There is a difference in the atheist population in US since 2005 to 2012 (Ha: p.1 != p.2)

us05 <- subset(atheism, nationality == "United States" & year == 2005)
us12 <- subset(atheism, nationality == "United States" & year == 2012)

table(us05$response)
table(us12$response)

inference(us05$response, success = "atheist", type = "ci", method = "theoretical", est = "proportion")

inference(us12$response, success = "atheist", type = "ci", method = "theoretical", est = "proportion")
```
```{r, echo=FALSE}
print("The confidence interval for the US in 2005 is (0.0038:0.0161) while the interval in 2012 is (0.0364:0.0634). The intervals do not overlap which means there was a change in the atheist population between 2005 and 2012. We can therefore reject the null hypothesis.")
```

 
2. If in fact there has been no change in the atheism index in the countries listed in Table 4, in how many of those countries would you expect to detect a change (at a significance level of 0.05) simply by chance?
  
# Hint: Look in the textbook index under Type 1 error.
```{r}
possible_change = 0.05 * 39
sprintf("A Type 1 error is where we reject the null hypothesis incorrectly. In this case, the atheist population would not have changed much or at all in these countries. Using a confidence interval of 95%% we can be sure the 5%% of the listed countries would have a statistical significant change in their index. An estimated %2.f countries would have a change in their atheist proportions.", possible_change)
```

  3. Suppose you're hired by the local government to estimate the proportion of residents that attend a religious service on a weekly basis. According to the guidelines, the estimate must have a margin of error no greater than 1% with 95% confidence. You have no idea what to expect for \(p\). How many people would you have to sample to ensure that you are within the guidelines?
```{r}
 # To find the sample size we also need to find the SE and before we can calculate for n
me = 0.01
z = 1.96

# where me = z * se
se = me/z

# We can then find p using the scenario that over 50% of the sample will or have attended
# 
p = 0.5
# me = z * se
# me >= z * sqrt(p * (1 - p)/n)
# we can then convert this to make n the subject
# sample size => n = z^2 * p * (1 - p)/me^2

n = z^2 * p*(1-p)/me^2

paste("Possible sample size: ", n)
```

