---
title: 'Factor Analysis'
author: "Akane Simpson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries + functions

Load all the libraries or functions that you will use to for the rest of the assignment. It is helpful to define your libraries and functions at the top of a report, so that others can know what they need for the report to compile correctly. 
```{r libaries}
##r chunk
library(psych)

```

## The Data
The data for this assignment is IMdB Movie Reviews. The first column in the sentiment of the review. The second column is the number of tokens in the review. The remaining columns are word frequencies (percent of the number of tokens) for most frequent 250 terms in the corpus. Make sure to only use column 3:252 for PCA. The goal of the current assignment is to explore the themes used in movie reviews and determine which themes are used in positive versus negative reviews. 


```{r thedata}
##r chunk
imdb_review <- read.csv('IMdB Reviews MEH.csv')
```

## How many factors?

- Explore how many factors you should use.
  - Include a scree plot.
- How many components would make sense based on the scree plot?
  - ANSWER:
    Based on dataframe we have a total of 252 columns where the first two are 'Sentiment' and 'RawTokenCount'. Since Sentiment will be the independent variable and the 'RawTokenCount' does not affect sentiment will use the other variables. After plotting the scree plot we can see that the optimal number of components is 8.

```{r howmany}
##r chunk
colnames(imdb_review)
comp1 <- principal(imdb_review[, 3:252], nfactors = 5, rotate = 'varimax')

plot(comp1$values[1:20])
```

## Simple structure - run the PCA

- Run a least two PCA solutions in R
  - Note: Given the sample size, should limit number of components to 10 at most.
  - Include the saved `principal` code, but then be sure to print out the results, so the summary is on your report.

```{r solution 1}
##r chunk
comp5 <- principal(imdb_review[, 3:252], nfactors = 5, rotate = 'varimax')
loadings5 <- as.data.frame(comp5$loadings[1:250, 1:5])
theme1_5 <- subset(loadings5[order(loadings5$RC1, decreasing = T),], RC1 >= 0.15, select = c(RC1))
theme2_5 <- subset(loadings5[order(loadings5$RC2, decreasing = T),], RC2 >= 0.15, select = c(RC2))
theme3_5 <- subset(loadings5[order(loadings5$RC3, decreasing = T),], RC3 >= 0.15, select = c(RC3))
theme4_5 <- subset(loadings5[order(loadings5$RC4, decreasing = T),], RC4 >= 0.15, select = c(RC4))
theme5_5 <- subset(loadings5[order(loadings5$RC5, decreasing = T),], RC5 >= 0.15, select = c(RC5))

comp5
```

```{r solution 2}
##r chunk
comp8 <- principal(imdb_review[, 3:252], nfactors = 8, rotate = 'varimax')
loadings8 <- as.data.frame(comp8$loadings[1:250, 1:8])
theme1_8 <- subset(loadings8[order(loadings8$RC1, decreasing = T),], RC1 >= 0.15, select = c(RC1))
theme2_8 <- subset(loadings8[order(loadings8$RC2, decreasing = T),], RC2 >= 0.15, select = c(RC2))
theme3_8 <- subset(loadings8[order(loadings8$RC3, decreasing = T),], RC3 >= 0.15, select = c(RC3))
theme4_8 <- subset(loadings8[order(loadings8$RC4, decreasing = T),], RC4 >= 0.15, select = c(RC4))
theme5_8 <- subset(loadings8[order(loadings8$RC5, decreasing = T),], RC5 >= 0.15, select = c(RC5))
theme6_8 <- subset(loadings8[order(loadings8$RC5, decreasing = T),], RC6 >= 0.15, select = c(RC6))
theme7_8 <- subset(loadings8[order(loadings8$RC5, decreasing = T),], RC7 >= 0.15, select = c(RC7))
theme8_8 <- subset(loadings8[order(loadings8$RC5, decreasing = T),], RC8 >= 0.15, select = c(RC8))

comp8
```


## Adequate solution

- Examine the fit indices of your two solutions. Which solution has better fit indices? 
  - ANSWER:
    Based on the PCA solutions we see that 'comp8' has the better fit indices for two of three indices measured.

```{r fit indices comp5, warning=FALSE}
fa.stats(imdb_review[,3:ncol(imdb_review)], comp5)$rms
fa.stats(imdb_review[,3:ncol(imdb_review)], comp5)$RMSEA
fa.stats(imdb_review[,3:ncol(imdb_review)], comp5)$TLI
```


```{r fit indices comp8, warning=FALSE}
fa.stats(imdb_review[,3:ncol(imdb_review)], comp8)$rms
fa.stats(imdb_review[,3:ncol(imdb_review)], comp8)$RMSEA
fa.stats(imdb_review[,3:ncol(imdb_review)], comp8)$TLI
```

## Themes

- Pick what you think is the best solution (based primarily on interpretability). Print the themes in the r chunk below. What labels would you give the themes? How would you describe each theme? Provide both a label for each theme and a longer description of what each seems to be capturing here.

  - ANSWER:
    Based on the indices for both solutions comp8 was selected. Although comp5 may have the better RMS the total indices measured between the two compositions indicates that comp8 has the better indices. 
    - Theme 1 Label : Unknown Theme; Description: Harsh critique on a film based on plot, actor choices, and script. The reviewer mostly complained about the film but spent little time talking about the film itself.
    - Theme 2 Label : Drama; Description: Reviewer enjoyed this film and praised every elements of it. However, did not mention any real details about the film itself
    - Theme 3 Label : Musical; Description: Very positive review of film. Reviewer takes note of many production characteristics of film and encourages others to view the film as well.
    - Theme 4 Label : Storybook; Description: Positive review on a previous story or real life event. Possible heavy focus on characters with some possible changes to invoke more emotions.
    - Theme 5 Label : TV-Series; Description: A TV series based on a book. Reviewer mentions either an original run of the same series. The TV series may have some noticeable production value and original ideas that may differ from the source. 
    - Theme 6 Label : Horror: Description: A low budget horror film with emphasis on special effects and less on acting. Reviewer sounds neutral in this film review.
    - Theme 7 Label : New Release; Description: A review on a just released production with a notable cinema experience. Reviewer is impressed over director's use of camera, sound and direction.
    - Theme 8 Label : Kid Series; Description: A kid-friendly TV show about a boy growing up in a possible new environment. Includes themes of family.


```{r theme 1}
# Label : Harsh Critique
# Theme 1 has a negative outlook. No definite theme was detected 
theme1_8
```

```{r theme 2}
# Label: Positive Critique
theme2_8
```

```{r theme 3}
theme3_8
```

```{r theme 4}
theme4_8
```

```{r theme 5}
theme5_8
```

```{r theme 6}
theme6_8
```

```{r theme 7}
theme7_8
```

```{r theme 8}
theme8_8
```


## Group Differences
- Create theme scores for your best solution. 
- Examine the differences between positive versus negative reviews. Include summary statistics and t-test results. Be sure to interpret the t-test results in the following question
- Which themes are used more in positive reviews? Which themes are used more in negative reviews?

  - ANSWER: From the summary statistics and t-test we see that all 8 themes are used more in positive reviews. This is somewhat odd as if all topics are linked to positive sentiment that would mean some of the themes created are incorrect in some way or most of the reviews are positive.

```{r group differences 1}
imdb_review$Theme1 = apply(imdb_review[,row.names(theme1_8)], MARGIN = 1, sum)
imdb_review$Theme2 = apply(imdb_review[,row.names(theme2_8)], MARGIN = 1, sum)
imdb_review$Theme3 = apply(imdb_review[,row.names(theme3_8)], MARGIN = 1, sum)
imdb_review$Theme4 = apply(imdb_review[,row.names(theme4_8)], MARGIN = 1, sum)
imdb_review$Theme5 = apply(imdb_review[,row.names(theme5_8)], MARGIN = 1, sum)
imdb_review$Theme6 = apply(imdb_review[,row.names(theme6_8)], MARGIN = 1, sum)
imdb_review$Theme7 = apply(imdb_review[,row.names(theme7_8)], MARGIN = 1, sum)
imdb_review$Theme8 = apply(imdb_review[,row.names(theme8_8)], MARGIN = 1, sum)
```


```{r group differences 2}
summary(imdb_review$Theme1)
wilcox.test(imdb_review$Theme1 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme2)
wilcox.test(imdb_review$Theme2 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme3)
wilcox.test(imdb_review$Theme3 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme4)
wilcox.test(imdb_review$Theme4 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme5)
wilcox.test(imdb_review$Theme5 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme6)
wilcox.test(imdb_review$Theme6 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme7)
wilcox.test(imdb_review$Theme7 ~ imdb_review$Sentiment)
```

```{r}
summary(imdb_review$Theme8)
wilcox.test(imdb_review$Theme8 ~ imdb_review$Sentiment)
```

## Discussion Questions

- Describe a set of texts and research question that interests you that could be explored using this method. 

  - ANSWER: One set of texts that may be of interest is examining the topics in various metal music related bands. The research question that could be proposed is what topics do these bands interweave within their songs within the metal genre. To do this, one could collect song lyrics from popular metal artists albums, singles and EP and create and create a word matrix then performing PCA test to select limited factors we can identify the themes and what bands have lyrics based on those themes. This research can be important as it is a way for avid music researchers can further interconnect other genres that express similar themes.