---
title: 'Topics Models'
author: "STUDENT NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this assignment, you will use the similar data as the Factor Analysis assignment to discover the topics in movie reviews. However, instead of a csv file with word frequencies, the data set for this assignment has only the raw text which you will need to process as shown in class. 

## Load the libraries + functions

Load all the libraries or functions that you will use to for the rest of the assignment. It is helpful to define your libraries and functions at the top of a report, so that others can know what they need for the report to compile correctly.

Load the Python libraries or functions that you will use for that section. 

```{r}
##r chunk
library(reticulate)

```

```{python}
##python chunk

```

## The Data

```{r data}
##r chunk

```

## Create the Topics Model

Create the corpus for the model in R. 

```{r}
##r chunk

```

Clean up the text and create the Document Term Matrix. 

```{r}
##r chunk 

```

Weight the matrix to remove all the high and low frequency words. 

```{r}
##r chunk

```

Run and LDA Fit model (only!).

```{r}
##r chunk

```

Create a plot of the top ten terms for each topic.

```{r}
##r chunk

```

Use dplyr to compare the use of the topics in positive versus negative sentiment reviews.

```{r}
##r chunk

```

## Gensim Modeling in Python

Transfer the df['Text'] to Python and convert it to a list for processing. 

```{python}
##python chunk

```

Process the text using Python. 

```{python}
##python chunk

```

Create the dictionary and term document matrix in Python.

```{python}
##python chunk

```

Create the LDA Topics model in Python using the same number of topics as used in the Factor Analysis assignment. 

```{python}
##python chunk

```

Create the interactive graphics `html` file. Please note that this file saves in the same folder as your markdown document, and you should upload the knitted file and the LDA visualization html file. 

```{python}
##python chunk

```

## Interpretation

Interpret your topics and compare to MEM themes with PCA. Explain the results from your analysis (at least 5 sentences). 
  
ANSWER:

## Discussion Question

Thinking of your current job or prospective career, propose a research project to address a problem or question in your industry that would use topic models. Spend time thinking about this and write roughly a paragraph describing the problem/question and how you would address it with text data and topic modeling.

- ANSWER: 

