---
title: "Lecture_6_Assingment_6"
author: "Akane Simpson"
date: "'r Sys.Date()"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r Setting up the Data,echo=FALSE,include=FALSE}
#First Get the population and sample of Ames community Area
  load("./ames.RData")

  population <- ames$Gr.Liv.Area
  samp <- sample(population, 60)
  
  sample(population)
  sample(samp)

  samp_mean <- rep(NA, 50)
  samp_sd <- rep(NA, 50)
  n <- 60
  
  # Now we're ready for the loop where we calculate the means and standard deviations of 50 random samples.
  for(i in 1:50){
    samp <- sample(population, n) # obtain a sample of size n = 60 from the population
    samp_mean[i] <- mean(samp)    # save sample mean in ith element of samp_mean
    samp_sd[i] <- sd(samp)        # save sample sd in ith element of samp_sd
  }

#Lastly, we construct the confidence intervals.

  lower_vector <- samp_mean - 1.96 * samp_sd / sqrt(n) 
  upper_vector <- samp_mean + 1.96 * samp_sd / sqrt(n)
```


# ON YOUR OWN

1. Using the following function (which was downloaded with the data set), plot all intervals. What proportion of your confidence intervals include the true population mean? Is this proportion exactly equal to the confidence level? If not, explain why.

```{r Q1 Interval for Population}
plot_ci(lower_vector, upper_vector, mean(population))
```

```{r Q1a Proportion}
prop <- 1 - (3/50)
```
```{r Q1b Explanation,echo=FALSE}
sprintf("The proportion of sample intervals that do not have the population mean within their interval is %g. This means that the proportion does not equal to actual interval. Confidence Intervals show a range of values that the a population parameter could fall within. The value parameter itself does not have to fall within the interval.", prop)
```



2. Pick a confidence level of your choosing, provided it is not 95%. What is the appropriate critical value?
```{r Q2 Finding Z}
a <- 0.10
z <- 1 - a/2
```

```{r,echo=FALSE}
sprintf("Using the confidence interval of 90%% we get achieve the critical value of %.2f.", qnorm(z))
```


3. Calculate 50 confidence intervals at the confidence level you chose in the previous question. You do not need to obtain new samples, simply calculate new intervals based on the sample means and standard deviations you have already collected. 
```{r New Confidence Intervals}
  new_lower_vector <- samp_mean - z * samp_sd / sqrt(n)
  new_upper_vector <- samp_mean + z * samp_sd / sqrt(n)

```

  a. Using the  plot_ci  function, plot all intervals and calculate the proportion of intervals that include the true population mean. How does this percentage compare to the confidence level selected for the intervals?
```{r}
  plot_ci(new_lower_vector,new_upper_vector, mean(population))
```

```{r,echo=FALSE}
new_prop <- 1 - (17/50)
sprintf("A proportion of %g sample intervals do not have the population mean fall within their intervals. Compared to the 95%% CI, sample intervals are less likely to contain the population mean using the 90%% confidence interval.", new_prop)
```


