---
title: "Correlation and CFA Test For Voice Project"
author: 
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(moments)
library(pgirmess)
library(pastecs)
library(compute.es)
library(Hmisc)
library(corrplot)
library(psych)
library(semTools)
library(lavaan)
library(GPArotation)

```


# Dataset: 
    DV(12)
    - NARS The Negative Attitudes Towards Robots Scale
    - MC   Manipulation Check
    - AGIS Age Group Identification Scale
    - MEC  Measurement of Ethos/Credibility
    - SP   Social Presence
    - RSP  Robot Server Perception
    - PU   Perceived Usefulness
    - PEU  Perceived Ease of Use
    - ATSR Attitude Towards Social Robot
    - AI   Adoption Intention
    - AIQS Adjusted Interaction Quality Scale
    - ASQS Adjusted Service Quality Scale
    - IA2  Voice_Type * AGIS
    - IA3  Voice_Gender * ATSR
    
    IV(2)
    - Voice_Gender
    - Voice_Type



## Explore the dataset:
```{r starting} 
df_voice = read_excel("VDatav4.xlsx")
summary(df_voice)
str(df_voice)
```

#Correlation between 12 DVs

## Preparation
```{r Preparation}
df_cor <- df_voice %>%
  select(NARS,MC,AGIS,MEC,SP,RSP,PU,PEU,ATSR,AI,AIQS,ASQS)
str(df_cor)
```
As shown in the results above, everything is numeric, good to go.


## Correlation Matrix
```{r Correlation Matrix}
corM <- cor(df_cor)
corM_round2 <- round(corM, digits = 2)
corM_round2
```

## Matrix Visulization:
```{r Matrix Visulization}
corrplot(corM_round2,
         type = "upper",
         number.cex = 0.7,
         addCoef.col = 'black',
         order = 'AOE')
```


## Correlation Test :
```{r Correlation Test}
ptest <- rcorr(corM)
round(ptest$P,3)
```

As shown in above:

Note: SP basically has high correlation with every DVs, however, no significancy at all.

NARS AND MC has a correlation at 0.05 with a significant P value = 0.02
NARS AND MEC has a correlation at 0.4 with a significant P value = 0.04
NARS AND PEU has a correlation at 0.05 with a significant P value = 0.01
NARS AND ATSR has a correlation at 0.12 with a significant P value = 0.03
NARS AND AI has a correlation at -0.04 with a significant P value = 0.01

MC AND MEC has a correlation at -0.08 with a significant P value = 0.00
MC AND RSP has a correlation at 0.58 with a significant P value = 0.01
MC AND PU has a correlation at 0.56 with a significant P value = 0.01
MC AND PEU has a correlation at 0.6 with a significant P value = 0.00
MC AND ATSR has a correlation at 0.54 with a significant P value = 0.01
MC AND AIQS has a correlation at 0.48 with a significant P value = 0.03
MC AND ASQS has a correlation at 0.55 with a significant P value = 0.01

AGIS AND MEC has a correlation at 0.01 with a significant P value = 0.02
AGIS AND RSP has a correlation at 0.61 with a significant P value = 0.01
AGIS AND PU has a correlation at 0.59 with a significant P value = 0.01
AGIS AND PEU has a correlation at 0.47 with a significant P value = 0.05
AGIS AND ATSR has a correlation at 0.52 with a significant P value = 0.03
AGIS AND ASQS has a correlation at 0.5 with a significant P value = 0.05


MEC AND PU has a correlation at 0.02 with a significant P value = 0.003
MEC AND PEU has a correlation at -0.04 with a significant P value = 0.002
MEC AND ATSR has a correlation at -0.04 with a significant P value = 0.001
MEC AND AI has a correlation at -0.2 with a significant P value = 0.000
MEC AND AIQS has a correlation at 0.01 with a significant P value = 0.007


RSP AND PU has a correlation at 0.77 with a significant P value = 0.000
RSP AND PEU has a correlation at 0.64 with a significant P value = 0.002
RSP AND ATSR has a correlation at 0.65 with a significant P value = 0.001
RSP AND AI has a correlation at 0.54 with a significant P value = 0.011
RSP AND AIQS has a correlation at 0.62 with a significant P value = 0.003
RSP AND ASQS has a correlation at 0.69 with a significant P value = 0.001

PU AND PEU has a correlation at 0.71 with a significant P value = 0.000
PU AND ATSR has a correlation at 0.71 with a significant P value = 0.000
PU AND AI has a correlation at 0.49 with a significant P value = 0.021
PU AND AIQS has a correlation at 0.66 with a significant P value = 0.002
PU AND ASQS has a correlation at 0.71 with a significant P value = 0.000

PEU AND ATSR has a correlation at 0.64 with a significant P value = 0.001
PEU AND AI has a correlation at 0.47 with a significant P value = 0.020
PEU AND AIQS has a correlation at 0.53 with a significant P value = 0.014
PEU AND ASQS has a correlation at 0.60 with a significant P value = 0.004

ATSR AND AI has a correlation at 0.51 with a significant P value = 0.010
ATSR AND AIQS has a correlation at 0.58 with a significant P value = 0.006
ATSR AND ASQS has a correlation at 0.63 with a significant P value = 0.002

AI AND AIQS has a correlation at 0.48 with a significant P value = 0.025
AI AND ASQS has a correlation at 0.47 with a significant P value = 0.026

AIQS AND ASQS has a correlation at 0.74 with a significant P value = 0.000






## Practice: 
```{r CFA Practice}
KMO(corM)
library(nFactors)
scree <- scree(df_cor)
print(scree)

```





## CFA Preparation: 
```{r CFA Preparation}
df_cfa <- df_voice %>%
  select(NARS,MC,AGIS,MEC,SP,RSP,PU,PEU,ATSR,AI,AIQS,ASQS,Voice_Gender,Voice_Type,VT,IA,IA2,IA3)
str(df_cfa)

```


## Regression
```{r Regression}
model_regression <- lm(AI ~ AGIS, df_cfa)
summary(model_regression)

model_regression1 <- lm(AI ~ AGIS * Voice_Type, df_cfa)
summary(model_regression1)

```


## Anova 
```{r Anova}
model_anova <- aov(AI ~ factor(AGIS) , df_cfa)
summary(model_anova)


model_anova1 <- aov(AI ~ factor(AGIS) * factor(Voice_Type) , df_cfa)
summary(model_anova1)
```


## CFA:
```{r CFA}
aaa<-'
ai1 =~ AI_Q1 + AI_Q2 + AI_Q3
agis1 =~ AGIS_Q1 + AGIS_Q2 + AGIS_Q3 + AGIS_Q4 + AGIS_Q5 + AGIS_Q6 + AGIS_Q7 + AGIS_Q8 + AGIS_Q9 + AGIS_Q10 + AGIS_Q11 + AGIS_Q12 + AGIS_Q13
sp1 =~ SP_Q1 + SP_Q2 + SP_Q4 + SP_Q5 + SP_Q6 + SP_Q7
pu1 =~ PU_Q1 + PU_Q2 + PU_Q3 + PU_Q4 + PU_Q5 
peu1 =~ PEU_Q1 + PEU_Q2 + PEU_Q3 + PEU_Q4 + PEU_Q5
atsr1 =~ ATSR_Q1 + ATSR_Q2 + ATSR_Q3'

model_cfa <- cfa(aaa,df_voice)
summary(model_cfa, fit.measures = TRUE, standardized = TRUE)

AI1 <- c("AI_Q1", "AI_Q2", "AI_Q3")
alpha.pa1 = alpha(df_voice[AI1])
print(alpha.pa1, digits = 3)


AGIS1 <- c("AGIS_Q1", "AGIS_Q2", "AGIS_Q3", "AGIS_Q4","AGIS_Q5", "AGIS_Q6", "AGIS_Q7", "AGIS_Q8", "AGIS_Q9","AGIS_Q10", "AGIS_Q11", "AGIS_Q12", "AGIS_Q13")
SP1 <- c("SP_Q1", "SP_Q2", "SP_Q3", "SP_Q4","SP_Q5", "SP_Q6", "SP_Q7")
PU1 <- c("PU_Q1", "PU_Q2", "PU_Q3", "PU_Q4","PU_Q5")
PEU1 <- c("PEU_Q1", "PEU_Q2", "PEU_Q3", "PEU_Q4","PEU_Q5")
ATSR1 <- c("ATSR_Q1", "ATSR_Q2", "ATSR_Q3")


alpha.pa1 = alpha(df_voice[AI1])
print(alpha.pa1, digits = 3)

alpha.pa1 = alpha(data[AGIS1])
print(alpha.pa1, digits = 3)

alpha.pa1 = alpha(data[SP1])
print(alpha.pa1, digits = 3)

alpha.pa1 = alpha(data[PU1])
print(alpha.pa1, digits = 3)

alpha.pa1 = alpha(data[PEU1])
print(alpha.pa1, digits = 3)

alpha.pa1 = alpha(data[ATSR1])
print(alpha.pa1, digits = 3)


```

