---
title: "ANLY 510 Exam 1 Q1"
author: "Akane Simpson"
date: "12/12/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(moments)
library(pastecs)
library(compute.es) 
```

# Question
The data below represents the yearly earnings (in $1000s of dollars) that high school and college (BS) graduates earn at a small firm. Determine if there is any difference in pay for the two groups: Be sure to include a complete analysis (i.e., all assumptions checked and full understanding of effects garnered) and a clear summary with all needed statistical details included (see lab keys for examples).

# R Code
```{r, eval=FALSE}
library(readxl)
library(moments)
library(psych)

Q1 = read_xlsx('~/Downloads/Exam1Q1.xlsx')

str(Q1)
summary(Q1)
describe(Q1)

# Skewness
plot(density(Q1$Highschool))
agostino.test(Q1$Highschool)

plot(density(Q1$BS))
agostino.test(Q1$BS)
# Skewness Assumption made for both

# Normality
qqnorm(Q1$Highschool)
shapiro.test(Q1$Highschool)

qqnorm(Q1$BS)
shapiro.test(Q1$BS)
# Normality assumption made

# Equal Variance
Q1.lm = lm(Highschool ~ BS, Q1)
Q1.res = resid(Q1.lm)

plot(Q1$Highschool, Q1.res)
abline(0,0)

plot(Q1$BS, Q1.res)
abline(0,0)
# Equal Variance assumption made.

# t test Analysis
t.test(Q1$Highschool, Q1$BS, var.equal = TRUE, alternative = 'two.sided')

```

# APA Summary

In the current study, the researcher examined the difference of pay between two levels of education (high school and Bachelor of Science). Performing an independent t-test (equal variances assumed) we find there is no significant difference between education levels of High School (M = 39.51; SD = 2.49) and Bachelor of Science (M = 50.77; SD = 6.00), t(76) = -10.812, p < 0.01.
