---
title: "ANLY 510 Exam 1 Q2"
author: Akane Simpson
date: 12/12/2021
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moments)
library(readxl)
library(pastecs) 
library(compute.es) 
```

# Exam Question
The data below represents the number of driving errors made by drivers who did or did not attend driving school (coded 1 if they didn’t, 2 if they did) and at day or night (coded 1 for day, 2 for night). Testing was performed on Saturdays and Sundays (coded 1 Saturday, 2 Sunday).

Determine if there is any effect of driving school and whether its effect is influenced by lighting conditions. Be sure to include a complete analysis (i.e., all assumptions checked, and full understanding of effects garnered) and a clear summary with all needed statistical details included.

From the problem we can identify the variables that are independent, dependent, conditions and blockers
- errorsmade (Dependent)
- attendeddrivingschool (Independent)
- dayornight (Condition)
- dayoftesting (Blocker)

# R Code
```{r, eval=FALSE}
Q2 = read_xlsx('~/Downloads/Exam1Q2.xlsx')

str(Q2)
summary(Q2)
Q2$attendeddrivingschool = factor(Q2$attendeddrivingschool)
Q2$dayornight = factor(Q2$dayornight)
Q2$dayoftesting = factor(Q2$dayoftesting)

# Skewness
plot(density(Q2$errorsmade))
agostino.test(Q2$errorsmade)
# p value < 0.05, skewness assumption not made

# Kurtosis
anscombe.test(Q2$errorsmade)
# p value > 0.05. Kurtosis assumption made

# Normality
qqnorm(Q2$errorsmade)
shapiro.test(Q2$errorsmade)
# p value < 0.05. Normality assumption made

# Independence
Q2.lm = lm(errorsmade ~ factor(dayornight),Q2)
Q2.resd = resid(Q2.lm)
plot(Q2$errorsmade, Q2.resd)
abline(0,0)
# Independence assumption made

# Equal Variance
bartlett.test(Q2$errorsmade, factor(Q2$attendeddrivingschool))
tapply(Q2$errorsmade, factor(Q2$dayornight), var)
#  Variance value is less than 3. Variance assumption made

# Additivity
Q2.add = aov(errorsmade ~ attendeddrivingschool*dayornight, Q2)
summary(Q2.add)
# There is a significant interaction to attendeddriving school with
# the dayornight variable.

Q2.block = aov(errorsmade ~ attendeddrivingschool*dayoftesting, Q2)
summary(Q2.block)
# There is no significant interaction to attendeddrivingschool with
# the dayornight variable.

# ANOVA Analysis
Q2.aov = aov(errorsmade ~ attendeddrivingschool, Q2)
Q2.aov2 = aov(errorsmade ~ attendeddrivingschool + dayornight, Q2)
Q2.aov3 = aov(errorsmade ~ attendeddrivingschool + dayoftesting, Q2)

anova(Q2.aov, Q2.aov2, Q2.aov3)
# From the results adding dayornight to attendeddrivingschool has a significant
# difference to the model. adding dayoftesting does not show a significant 
# difference but does explain some of the variance.

summary(Q2.aov2)

interaction.plot(Q2$attendeddrivingschool, Q2$dayornight, Q2$errorsmade)

# Post Hoc Analysis
TukeyHSD(Q2.aov2)

# Effect Size
by(Q2$errorsmade, Q2$attendeddrivingschool, stat.desc)

mes(0.202, 0.283, 0.07, 0.12, 24, 24)
```

# APA Summary

A study was conducted to investigate whether, the average errors made by a student on a driving exam  taken together, if the student had attended a driving school before, the time of day affected individuals’ overall errors made. Observations from the study were analyzed by conducting a two-way analysis of variance with the three independent variables (attendeddrivingschool, dayornight or dayoftesting) and the dependent variable (errorsmade) using R version 4.12.

First, all assumptions are met and there is no adjustment made. ANOVA analysis revealed there is statistically significant interaction effect (F(1, 44) = 4.733, p = .035). Results also suggest that the errors made on the driving exam were affected by the time of day of (day or night) the participants (F(1, 44) = 22.481, p < .001) and between attending and not attending driving school (F(1, 44) = 13.146, p < .001).

The effect size was large D = 0.82 between attending and not attending driving school.