---
title: "ANLY 510 Assignment 3"
author: "Akane Simpson"
date: "11/27/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

The data represents a study of older participants that tests their fine motor skill performance. The participants are split into three age groups. From there participants are given a random task and are graded on their performance.


## Hypothesis

  From the research question we can establish a null and alternative hypothesis in regards to the motor skille performance. 
  
  - $H_O$: There is no significant difference in motor skills between the three groups. 
  - $H_A$: There is at least one bage group that has a signifcant difference in motor skills performance.
  
## Assumptions

  From the research we were able to identify three variable
  
  - Performance Score
  - Condition
  - Age (*group “1”:60-69, “2”: 70-79 and “3”: above 80)
  
### Specification Table
  
  |__Variable__   |__# Levels__   |
  |:---------:|:---------:|
  |Age        |3          |
  |Condition  |3          |
  
Where out independent variable (IV) is Age while our dependent variable (DV) is Performance Score. This leaves our blocker (Blocker) as Condition. We can then make a table to represent the relationship between the IV and Blocker. 
  
  |Block/IV   |Test 1     |Test 2     | Test 3      |
  |:---------:|:---------:|:---------:|:-----------:|
  |Age 1      |10         |10         |10           |
  |Age 2      |10         |10         |10           |
  |Age 3      |9          |10         |10           |

## R Code

```{r rcode, eval=FALSE}
library(moments)
library(readxl)
library(pgirmess)
library(pastecs)
library(compute.es)

master = readxl::read_xlsx('./Week 3/Lab3.xlsx')
data = master

summary(data)

# Skewness
plot(density(data$Performance_score))
agostino.test(data$Performance_score)
# Skewness assumption is not broken

# Kurtosis
anscombe.test(data$Performance_score)
# Anscombe test shows high kurtosis of data. 
# Kurtosis assumption is broken with p < 0.05

# Normality
qqnorm(data$Performance_score)
shapiro.test(data$Performance_score)
# Shapiro test and Q-Q plot shows points points are mostly normal. p-value < 0,05
# Normality assumption is not broken.

# Independence
model.lm = lm(Performance_score ~ Age , data = data)
model.res = resid(model.lm)
plot(data$Performance_score, model.res,)
abline(0,0)
# Plot shows values being evenly distributed above and below abline. 
# Independence assumption is not broken

# Equal Variance
bartlett.test(data$Performance_score, data$Age)
tapply(data$Performance_score, data$Age, var)
# Results show there is no major variance across the levels and p > 0,05. Equal Variance
# assumption is not broken

# Additivity of Interaction
model = aov(Performance_score ~ factor(Age) * factor(Condition), data = data)
summary(model.aov)
# The Interaction model 1 shows that there is a  significant interaction between
# Condition and Age. Residuals show that Condition and its connection does explain
# some of the error in the model.

# ANOVA Analysis
model.aov2 = aov(Performance_score ~ factor(Age) + factor(Condition), data = data)
summary(model.aov2)

model.aov3 = aov(Performance_score ~ factor(Age), data = data)
summary(model.aov3)

anova(model.aov2, model.aov3)
# Comparing the model with the blocker versus without the blocker shows that
# condition (blocker) does have an effect on the model's residuals and explains
# some of the variance.


# Post Hoc Analysis
pairwise.t.test(data$Performance_score, data$Age, p.adjust.method = "bon", 
                paired = FALSE)
# Bonferroni test shows there is a significant difference between each age groups
# in terns of motor skills performance.

# Effect Size
by(data$Performance_score, data$Age, stat.desc)

# Groups 2 and 1
mes(27.94, 32.45, 4.35, 3.59, 31, 29)

# Group 3 and 1
mes(22.14, 32.45, 3.98, 3.59, 29, 29)

# Group 3 and 2
mes(22.14, 27.94, 3.98, 3.59, 29, 31)

# Using mes we compare the difference between age groups to find the effect size
```

## APA Summary

Observations from the study were analyzed by conducting a one-way analysis of variance using R version 4.1.1. First, all assumptions are met and there is no adjustment made. Results suggest that the performance level, after controlling for condition, between age groups (as 1, 2 or 3) of the participants (F(2, 86) = 48.48, p < .001).  
  
Continue the discussion with specifically which groups differed, a Bonferroni test was established. The result suggested that there is a significant difference between age groups 1 and 2 (p < .001) and age groups 1 and 3 (p < .001), in terms of the perfromance levels. There was also a significant difference between age groups 2 and 3. The effect was large between methods 1 & 2, and 1 & 3, Cohen’s D = 1.13 and 2.72. The effect size for groups 2 and 3 was Cohen's D = 1.53.  
