---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# import data 

```{r}
NY_death = read.csv('./NY_death.csv')

dim(NY_death)
```

#convert date


#
```{r}
NY_death$date = as.Date(NY_death$date, format = "%Y-%m-%d")

head(NY_death$date)

summary(NY_death$date)
```

# selected columns


```{r}
cols = c('Male_age65plus', 'ICU.Beds', 'Median_Household_Income_2018', 'Total.Hospitals..2019.', 'Employed_2018', 
         'Unemployed_2018', 'Percent.of.adults.with.less.than.a.high.school.diploma.2014.18',
       "Percent.of.adults.with.a.bachelor.s.degree.or.higher.2014.18", 'confirmed', 'Housing.units',
     'Cardiovascular.Disease..AAMC.', 'Pulmonary.Disease..AAMC.')
```

# size of the key columns

```{r}
length(cols)
```


# import libraries

```{r}
library(GGally)
library(reshape2)
library(ggplot2)
library(grid)

library(zoo)
```

# clean the data

```{r}

print('before fixing the NaN values ...........................................................')
apply(NY_death, 2, function(x){sum(is.na(x))}) # check the missing column

# use forward fill for NaN
NY_death = na.locf(na.locf(NY_death), fromLast = TRUE)


print('after fixing the NaN values ...........................................................')
apply(NY_death, 2, function(x){sum(is.na(x))}) # check the missing column
```


# remove the outliers

```{r}


NY_death_2020 = subset(NY_death, format.Date(date, "%Y")=="2020")
NY_death_2021 = subset(NY_death, format.Date(date, "%Y")=="2021")

# calculate mahal distance 2020
mahal_score2020 = mahalanobis(NY_death_2020[, cols],
         colMeans(NY_death_2020[, cols], na.rm=TRUE),
         cov(NY_death_2020[, cols]), use="pairwise.complete.obs",tol=1e-26)


summary(mahal_score2020)

cutoff_2020 = qchisq(0.999, ncol(NY_death_2020[, cols]))

print(cutoff_2020)

NY_death_2020_new = subset(NY_death_2020, mahal_score2020 < cutoff_2020)




mahal_score2021 = mahalanobis(NY_death_2021[, cols],
         colMeans(NY_death_2021[, cols], na.rm=TRUE),
         cov(NY_death_2021[, cols]), use="pairwise.complete.obs",tol=1e-26)


# summary(mahal_score2021)

cutoff_2021 = qchisq(0.999, ncol(NY_death_2021[, cols]))

print(cutoff_2021)

NY_death_2021_new = subset(NY_death_2021, mahal_score2021 < cutoff_2021)
```


# check for linearity

```{r}
dim(NY_death_2020_new[, cols])
random_data = rchisq(nrow(NY_death_2020_new[, cols]), 7)
#fake regression
fake_data2020 = lm(random_data~., data = NY_death_2020_new[, cols])

standardized2020 = rstudent(fake_data2020)
fit_value2020 = scale(fake_data2020$fitted.values)

plot(fake_data2020, 2, main = 'data QQ plot: 2020')


dim(NY_death_2021_new[, cols])
random_data = rchisq(nrow(NY_death_2021_new[, cols]), 7)
#fake regression
fake_data2021 = lm(random_data~., data = NY_death_2021_new[, cols])

standardized2021 = rstudent(fake_data2021)
fit_value2021 = scale(fake_data2021$fitted.values)

plot(fake_data2021, 2, main = 'data QQ plot: 2021')


print('overall the distribution look fine. There is a slight positive skewness')

```

# check normality

```{r}
print('hist 2020')
hist(standardized2020, breaks = 100)
print('hist 2021')
hist(standardized2021, breaks = 100)

library(moments)
print('skewness check for 2020')
skewness(NY_death_2020_new[, cols],na.rm=TRUE)
kurtosis(NY_death_2020_new[, cols],na.rm=TRUE)-3


print('skewness check for 2021')
skewness(NY_death_2021_new[, cols],na.rm=TRUE)
kurtosis(NY_death_2021_new[, cols],na.rm=TRUE)-3
```


# data screening: check homogeneity/homoscedasticity 
```{r}
{plot(fit_value2020, standardized2020)
abline(0,0)
abline(v = 0)}

plot(fake_data2020,1)
```


# homogeneity 2021
```{r}
{plot(fit_value2021, standardized2021)
abline(0,0)
abline(v = 0)}

plot(fake_data2021,1)
```



# corrplot 2020 
```{r}
library(corrplot)
print('2020 corrplot')


#png('corrplot_2020.png') # dump out to disk

corrplot(cor(NY_death_2020_new[, cols]), type = "upper", tl.pos = "td",
         method = "circle", tl.cex = 0.5, tl.col = 'black',
         order = "hclust", diag = FALSE)
#dev.off() # dump out to disk

```


# corrplot 2021
```{r}
print('2021 corrplot')


#png('corrplot_2021.png')

corrplot(cor(NY_death_2021_new[, cols]), type = "upper", tl.pos = "td",
         method = "circle", tl.cex = 0.5, tl.col = 'black',
         order = "hclust", diag = FALSE)
#dev.off() # dump out disk
```



# define the custom plot function and identify the months



```{r}

all_months = c('03', '04', '05', '06', '07', '08', '09')

#NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
#NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")
  
# reference:  https://stackoverflow.com/questions/5861489/r-creating-legend-for-three-data-sets-on-same-graph-using-ggplot  
my_plot = function(NY_death2020_M, NY_death2021_M, current_col){
  NY_death2020_M = unique(NY_death2020_M[, c(current_col, 'death_count')])
  NY_death2021_M = unique(NY_death2021_M[, c(current_col, 'death_count')])
  
  x = NY_death2020_M[, current_col]
  y = NY_death2020_M$death_count
  
  x3 = NY_death2021_M[, current_col]
  y3 = NY_death2021_M[, current_col]
  
  p1 = data.frame(x = x, y = y)
  p3 = data.frame(x = x3, y = y3)
  
  zz = melt(list(p20 = p1, p21 = p3), id.vars = 'x')
  
  ggplot(zz, aes(x, value, color=L1)) + geom_point() +  ggtitle(paste0(current_col)) +
    scale_colour_manual("Data", values = c("p20" = "blue", "p21" = "red")) 
}


```

# define custom plot function


```{r}


my_multiplot = function(NY_death2020_M, NY_death2021_M, cols, month){
  grid.newpage()
  
  #pdf_file_name = paste0('month_', str(month), '.pdf')
  #pdf(pdf_file_name, height = 11, width = 8.5, paper = "letter")
  row_size = 3
  
  pushViewport(viewport(layout = grid.layout(4,row_size)))
  vplayout <- function(x,y){viewport(layout.pos.row = x, layout.pos.col = y)}
  
  row_index = 1
  
  for (count in seq(1, row_size * 4)){
    col_index = count %% row_size
    if (col_index == 0){
      col_index = row_size
    }
    
    
    
    my_curr_plot = my_plot(NY_death2020_M, NY_death2021_M, cols[count])
    print(my_curr_plot, vp = vplayout(row_index, col_index))
  
    if (col_index == row_size){
      row_index = row_index + 1
    }
  
  }
}

```

# define the models generation function
```{r}
# Reference https://stackoverflow.com/questions/17024685/how-to-use-reference-variables-by-character-string-in-a-formula

dfmodels.analysis <- function(df1, df2, type = '') {
  current_month = months(df1$date[1], abbreviate = T)
  frm <- 'death_count ~ confirmed'

  if(type == 'environmental') {
    print('Environment Selected')  
    # browser()
    
    # Using some of the environment variables
    # Precipitation, Average, Min & Max Temperature
    precip <- paste(current_month, '.Precipitation...inch',sep = '')
    avg.temp <- paste(current_month, '.Temp.AVG...F', sep='')
    min.temp <- paste(current_month, '.Temp.Min...F', sep = '')
    max.temp <- paste(current_month, '.Temp.Max...F', sep='')
    
    predictors <- paste('date',
                        precip,
                        avg.temp,
                        min.temp,
                        max.temp,
                        sep = ' + ')
    frm <- as.formula(paste(frm, predictors, sep = ' + '))
  }
  
  else if(type == 'social') {
    print('Social Selected')
    # browser()
    # Using some of the social and economical variables
    # Population, 
    # Total_Male, 
    # Total_Female
    # Area.in.square.miles...Total.area
    # Density.per.square.mile.of.land.area...Population
    # Civilian_labor_force_2018
    # Total_households...Average.household.size
    # Total_households...Average.family.size
    
    # Need to add another predictor? Just add a string in the paste function of predictors!
    predictors <-  paste(
             'date',
             'Population',
             'Density.per.square.mile.of.land.area...Population',
             'Area.in.square.miles...Total.area',
             'Total.households',
             'Total.households..Average.household.size',
             'Total.households..Average.family.size',
             # '',
             sep = ' + ')
    
    frm <- as.formula(paste(frm, predictors, sep = ' + '))
  }
  
  # TODO: Need to find and complete healthcare related predictors
  # else if(type == 'healthcare') {
  #   print('Healthcare Selected')
  #   predictors <- paste('', sep = ' + ')
  #   frm <- as.formula(paste(frm, predictors, sep = ' + '))
  # }
  else {
    print('No types selected. Defaulting to death_count ~ confirmed')
    frm <- as.formula(paste(frm, 'date', sep = ' + '))
  }


  # set.seed(3)
  
  # Sample from the data sets
  # sample1 <- df1[sample(nrow(df1), 100, replace = FALSE), ]
  # sample2 <- df2[sample(nrow(df2), 100, replace = FALSE), ]
  
  # Then create models from the samples
  model1 <- lm(frm, df1)
  model2 <- lm(frm, df2)

  print('DF Model 1 ')
  print(model1)
  print(summary(model1, correlation = T))
  
  print('DF Model 2')
  print(model2)
  print(summary(model2, correlation = T))

  # browser()

  # Let's compare with ANOVA
  # print(anova(model1, model2))
  return(list(model1,model2))
}
```


# draw multiplot on a per month basis


```{r}
month = '02'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")


#NY_death2020_M = read.csv('NY_death_2020_Month_2.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_2.csv')
```


```{r}
my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)

```

# linear model
```{r base model}
base <- dfmodels.analysis(NY_death2020_M, NY_death2021_M)
```


```{r env}
env <- dfmodels.analysis(NY_death2020_M, NY_death2021_M, 'environmental')
```


```{r social}
social <- dfmodels.analysis(NY_death2020_M, NY_death2021_M, 'social')
```

```{r prediction}
# sample(nrow(df1), 100, replace = FALSE)
# We get the first 3 days of the month in out old 
test_days <-  subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020" & format.Date(date, '%d') < '04')
pred.base.mon2 <- predict(base[[2]], test_days)

plot(test_days$date, test_days$death_count)
abline(pred.base.mon2)
```



# draw multiplot on a per month basis


```{r}
month = '03'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_3.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_3.csv')

my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```

# linear model
```{r}
dfmodels.analysis(NY_death2020_M, NY_death_2021)
```


```{r}
dfmodels.analysis(NY_death2020_M, NY_death_2021, "environmental")
```


```{r}
dfmodels.analysis(NY_death2020_M, NY_death_2021, "social")
```

```{r prediciton}

```


# draw multiplot on a per month basis


```{r}
month = '04'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_4.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_4.csv')

my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```
# linear model
```{r}

```



# draw multiplot on a per month basis


```{r}
dim(unique(NY_death2020_M))
dim(unique(NY_death2021_M))

# build the model to utilize all columns

current_col = cols[1]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[2]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[3]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[4]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')


current_col = cols[5]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[6]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[7]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')



current_col = cols[8]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[9]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[10]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')


current_col = cols[11]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')

current_col = cols[12]

print(current_col)
current_df = NY_death2020_M[, c(current_col, 'death_count')]
current_df = unique(current_df)
current_df = na.omit(current_df)

plot(current_df[, current_col], current_df$death_count, main = paste0(current_col))

curr_model = lm(current_df$death_count ~ current_df[, current_col])
summary(curr_model)
print('-------------------------------------------')






```
# linear model



# draw multiplot on a per month basis


# 
```{r}
month = '05'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_5.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_5.csv')

my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```

# linear model



# draw multiplot on a per month basis


```{r}
month = '06'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_6.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_6.csv')

my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```

# linear model



# draw multiplot on a per month basis

```{r}
month = '07'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_7.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_7.csv')

my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```


# linear model

# draw multiplot on a per month basis

```{r}
month = '08'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_8.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_8.csv')

my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```

# linear model



# draw multiplot on a per month basis


```{r}
month = '09'
NY_death2020_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2020")
NY_death2021_M = subset(NY_death, format.Date(date, "%m")==month & format.Date(date, "%Y")=="2021")

#NY_death2020_M = read.csv('NY_death_2020_Month_9.csv')
#NY_death2021_M = read.csv('NY_death_2021_Month_9.csv')
```


```{r}
my_multiplot(NY_death2020_M, NY_death2021_M, cols, month)
```
#
# linear model
```{r}

```

