---
title: "Module 3 - Fitting Discrete Distributions 2"
author: "Akane Simpson"
date: '2022-04-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(kableExtra)
library(vcd)
```

```{r}
base <- read.csv('./data.3.csv')
dice_groups <- ncol(base)
colNames <- names(base)
maxVals <- max(base, na.rm = TRUE)
nobs <- nrow(base)
```

```{r}
s123 <- rowSums(base == 1 | base == 2 | base == 3)
fs123 <- factor(s123, levels = 0:dice_groups)
freq.values <- xtabs(~fs123)
```

```{r}
freq.matrix <- matrix(data = freq.values, nrow = 1, byrow = TRUE, dimnames = list(c("Frequency"), c(0:dice_groups)))
table <- kable(freq.matrix, caption = "Number of successes when rolling a 8 sided dice in groups of 9. ") %>%
  kable_styling()
```

```{r, include=FALSE}
# Assuming p = 0.05 test if be fitted data from goodness-of-fit values shows the dice distribution is binomial

# H_0: The dice distribution is binomial (fit == binomial)
# H_A: The dice distribution is not binomial (fit != binomial)

fit <- goodfit(freq.values, type = "binomial")
fit.summa <- summary(fit)

# Since 0.894 > p then we should not reject the null
```

## Analysis (Fact)

Perform a goodness-of-fit test.

1. The test’s __statistic__ has a value of $x^{2}$ = __`r format(fit.summa[1], digits = 3)`__.
2. The test has a __p-value__ of __`r format(fit.summa[3], digits = 3)`__.
3. The test provides __evidence__ that the distribution [is] ___a binomial distribution___ because ___the returned p-value is greater than the $x^{2}$ value. Hence we must not reject the null hypothesis___.

## Tables and Figures

Find the maximum dice size: __`r maxVals`__

Find the group size: __`r dice_groups`__

Considering a roll of 1/2/3 to be a success, find the number of successes per grouping.

`r table`

```{r, warning=FALSE}
plot(x = fit, y = NULL, ylab = "Frequency", scale="raw", shade = T, main = "Number of successes when rolling a 8 sided dice.",   xlab = "Number of Successes")
```

