---
title: "Module 3 - Fitting Discrete Distributions 1"
author: "Akane Simpson"
date: '2022-04-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
```


```{r Data Wrangling}
base <- read.csv('./data.3.csv')
dice_groups <- ncol(base)
colNames <- names(base)
maxVals <- max(base, na.rm = TRUE)
nobs <- nrow(base)
```

## Analysis (Intuition)

1. The dice roll is believed to be [fair] because _every side of the dice has a 1/8 chance of being rolled_.
2. The distribution observed in the plot appears to be a _binomial_ distribution because _it is right skewed_.

## Tables and Figures

Find the maximum dice size: __`r maxVals`__

Find the group size: __`r dice_groups`__


Considering a roll of 1/2/3 to be a success, find the number of successes per grouping.

```{r}
success_fail_freq <- array(0:dice_groups+1)

for(row in 1:nobs){
  count <- 0
  
  for (val in base[row, ]) {
     if(val == 1 | val == 2 | val == 3){
       count <- count + 1
     }
  }
  success_fail_freq[count+1] = success_fail_freq[count+1] + 1
}

freq.matrix <- matrix(data = success_fail_freq, nrow = 1, byrow = TRUE, dimnames = list(c("Frequency"), c(0:dice_groups)))

table <- kable(freq.matrix, caption = "Number of successes when rolling a 8 sided dice in groups of 9. ") %>%
  kable_styling()
```

`r table`

```{r}
plot(c(0:dice_groups), freq.matrix[1,], xlab = "Number of Successes", ylab = "Frequency", main = "Number of successes when rolling a 8 sided dice in groups of 9")
```
