---
title: 'Similarity Assignment'
author: "STUDENT NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Python Application

Import the `completed_clean_data` and convert to a `pandas` dataframe. This dataset includes a list of scientific research articles that all appeared when I searched for "databases", "corpus", and "linguistic norms". 

```{r}

```


```{python}
##python chunk

```

Load all the libraries you will need for the Python section. You can also put in the functions for normalizing the text and calculating the top 5 related objects.

```{python}
##python chunk

```

Use the normalizing text function to clean up the corpus - specifically, focus on the `ABSTRACT` column as our text to match.

```{python}
##python chunk

```

Calculate the cosine similarity between the abstracts of the attached documents. 

```{python}
##python chunk

```

Using our moving recommender - pick a single article (under `TITLE`) and recommend five other related articles.

```{python}
##python chunk

```

## Make a Change to the Model

Using the methods shown in class, make one change to the model to see how it impacts the outcome. Pick one of the following: use a different similarity metric, use phrases instead of single words (e.g. change ngram_range), use only more frequent terms (e.g. change min_df), or lemmatize the words in the processing step.

```{python}
##python chunk

```


## Discussion Questions

- Did you get the articles expected? Do the suggestions make sense? Did your change to the model improve the recommendations? What else might improve the recommendation algorithm? 

 - ANSWER: 

- Describe a set of texts and research question that interests you that could be explored using this method. Basically, what is a potential application of this method to another area of research?

  - ANSWER: 