---
title: "Module 6 - Logistic Regression 2"
author: "Akane Simpson"
date: '2022-07-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(lmtest)
library(ggplot2)
```

```{r, results='hide'}
data <- read.csv('./data.6.csv')
data$Sex <- factor(data$Sex, labels = c("Male", "Female"))
data$Improved <- as.numeric(data$Improved)
colnames(data)[3] <- "Height"
colnames(data)[4] <- "Weight"
```

```{r, results='hide'}
sex.glm <- glm(Improved ~ Sex, data = data, family = "binomial")
# H0: Sex is not a significant factor of Improved
# HA: Sex is a significant factor of Improved
summary((sex.glm))
exp(coef(sex.glm))
```

```{r, results='hide'}
# Reverse Sex for Male
data$Sex<- factor(data$Sex, levels = rev(levels(data$Sex)))
sex.glm <- glm(Improved ~ Sex, data = data, family = "binomial")
summary((sex.glm))
exp(coef(sex.glm))
```

```{r, results='hide'}
# Weight_Sex
weight.glm <- glm(Improved ~ Weight, data = data, family = "binomial")
# H0: Weight is not a significant factor of Improved
# HA: Weight is a significant factor of Improved
summary(weight.glm)
exp(coef(weight.glm))
```

```{r, results='hide'}
# Height
height.glm <- glm(Improved ~ Height, data = data, family = "binomial")
# H0: Height is not a significant factor of Improved
# HA: Height is a significant factor of Improved
summary(height.glm)
exp(coef(height.glm))
```

```{r, results='hide'}
# Height_Weight interaction
height_weight.glm <- glm(Improved ~ Weight : Height, data = data, family = "binomial")
summary(height_weight.glm)
exp(coef(height_weight.glm))
```

```{r, results='hide'}
# Full combination
full_combination <- glm(Improved ~ Height + Weight + Sex, data = data, family = "binomial")
summary(full_combination)
exp(coef(full_combination))
```

```{r, results='hide'}
anova(sex.glm, height.glm, weight.glm, height_weight.glm, test = "Chisq")
```

## Analysis

1. Sex ___[is]___ a significant predictor of Improved (β = ___1.0845___, p = ___1e-05___).
2. Males show an ___[decrease]___ in the Odds of the subject’s symptoms by ___`r abs(0.338 - 1) * 100`___ %.
3. When looking at Weight, Height, Sex, and the interaction of Weight and Height at the same time, the below terms are significant (delete terms that are not significant).

    - Weight (β = ___0.06021___, p = ___2.33e-09___)
    - Sex (β = ___1.0845___, p = ___1e-05___)
    - Weight : Height (β = ___2.215e-04___, p = ___3.57e-06___)

4. When comparing the two models from last week and the two from this week on AIC, the ___[Weight only]___ model is the best.

## Table and Figures
```{r}
ggplot(data = data, aes(y = Improved, x = Height, color=Sex)) +
  geom_point(position = position_jitter(height = 0.02, width = 0, seed = 0)) +
  stat_smooth(formula = y ~ x, method = "glm", method.args = list(family="binomial"),
              alpha = 0.1, fill = 'blue', fullrange = TRUE) +
  theme_bw()

ggplot(data = data, aes(y = Improved, x = Weight, color=Sex)) +
  geom_point(position = position_jitter(height = 0.02, width = 0, seed = 0)) +
  stat_smooth(formula = y ~ x, method = "glm", method.args = list(family="binomial"),
              alpha = 0.1, fill = 'blue', fullrange = TRUE) +
  theme_bw()
```

