---
title: 'Word2Vec'
author: "Akane Simpson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries + functions

Load all the libraries or functions that you will use to for the rest of the assignment. It is helpful to define your libraries and functions at the top of a report, so that others can know what they need for the report to compile correctly.

```{r libaries}
##r chunk
library(reticulate)
```

Load the Python libraries or functions that you will use for that section. 

```{python}
##python chunk
import pandas as pd
import numpy as np
from bs4 import BeautifulSoup
import html
import re
import nltk
from gensim.models.fasttext import FastText
from nltk.corpus import stopwords
from sklearn.metrics import classification_report
from sklearn.metrics import accuracy_score
from sklearn.model_selection import train_test_split
from sklearn.linear_model import LogisticRegression

REPLACE_BY_SPACE_RE = re.compile('[/(){}\[\]\|@,;]') #remove symbols with space
BAD_SYMBOLS_RE = re.compile('[^0-9a-z #+_]') #take out symbols altogether
STOPWORDS = set(stopwords.words('english')) #stopwords

def clean_text(text):
    text = BeautifulSoup(text, "html").text # HTML decoding
    text = text.lower() # lowercase text
    text = REPLACE_BY_SPACE_RE.sub(' ', text) # replace REPLACE_BY_SPACE_RE symbols by space in text
    text = BAD_SYMBOLS_RE.sub('', text) # delete symbols which are in BAD_SYMBOLS_RE from text
    text = ' '.join(word for word in text.split() if word not in STOPWORDS) # delete stopwords from text
    return text

def document_vectorizer(corpus, model, num_features):
    vocabulary = set(model.wv.index_to_key)
    
    def average_word_vectors(words, model, vocabulary, num_features):
        feature_vector = np.zeros((num_features,), dtype="float64")
        nwords = 0.
        
        for word in words:
            if word in vocabulary: 
                nwords = nwords + 1.
                feature_vector = np.add(feature_vector, model.wv[word])
        if nwords:
            feature_vector = np.divide(feature_vector, nwords)

        return feature_vector

    features = [average_word_vectors(tokenized_sentence, model, vocabulary, num_features)
                    for tokenized_sentence in corpus]
    return np.array(features)
```

## The Data

The dataset is a set of Youtube comments that have been coded as:
  - 1: spam youtube messages 
  - 0: good youtube messages
  - This data is stored in the `CLASS` column

Import the data using either R or Python. I put a Python chunk here because you will need one to import the data, but if you want to first import into R, that's fine. 

```{python}
##python chunk
df = pd.read_csv('youtube_spam.csv')
df["CONTENT"].head()
```

## Clean up the data

Use one of our clean text functions to clean up the `CONTENT` column in the dataset. 

```{python}
##python chunk
df["CONTENT"] = df["CONTENT"].apply(clean_text)
df["CONTENT"].head()
```

## Split the data

Split the data into testing and training data.

```{python}
##python chunk
X = df["CONTENT"]
Y = df["CLASS"]

x_train, x_test, y_train, y_test = train_test_split(X, Y, test_size = 0.20, random_state = 876)
```

## Process the data

For FastText OR word2vec, create the tokenized vectors of the text.

```{python}
##python chunk
tokenized_train = [nltk.tokenize.word_tokenize(text) for text in x_train.to_list()]
tokenized_test = [nltk.tokenize.word_tokenize(text) for text in x_test.to_list()]
```

## Build Model

Build either a word2vec or FastText model.

```{python}
##python chunk
ft_model = FastText(tokenized_train, vector_size = 100, window = 6, min_count = 2, epochs = 5, seed = 876,  workers = 4)
```

## Convert the model

Convert the model into a set of features to use in our classifier. 

```{python}
##python chunk
avg_ft_train_features = document_vectorizer(corpus=tokenized_train, model=ft_model,
                                                     num_features=100)
avg_ft_test_features = document_vectorizer(corpus=tokenized_test, model=ft_model,
                                                    num_features=100)    
```

## Build a classifier model

In class, we used logistic regression to classify the data. You can use any machine learning algorithm you want here, and build a classification model. 

```{python}
##python chunk
logreg = LogisticRegression(solver='lbfgs', multi_class='ovr', max_iter=10000)
logreg = logreg.fit(avg_ft_train_features, y_train)
y_pred = logreg.predict(avg_ft_test_features)
```

## Examine the results

Print out the accuracy, recall, and precision of your model.

```{python}
##python chunk
print('accuracy %s' % accuracy_score(y_pred, y_test))
print(classification_report(y_test, y_pred,target_names=["good", "spam"]))
```

## Discussion Questions

- How well did the model you build do? What could you do to improve the accuracy of the model? (Interpret the model in at least 5 sentences)

  - ANSWER: The model performed moderately well in this analysis. The measured accuracy of this FastText model was 64.5% it scored slightly well on recall for good comments (55%) and scored well on spam comments (73%). The average precision and F1 score for both outcomes was 64%. To improve this model we could resampling the training and testing data, cleaning text further by including lemmatization or by changing the number of features used in our document vectorization. Additionally, a change in model (FastText -> Word2Vec) and altering the classifier could help improve the model performance. 

- Describe a set of texts and research question that interests you that could be explored using this method. Basically, what is a potential application of this method to another area of research?

  - ANSWER: One set of texts that would use as research is the examining genres in popular video games. A research question one could pose is whether the tastes and needs of American gamers have changed over the past decade. To do this we could create a list of popular video games (domestic and foreign) and tabulate their user rating, music, gameplay, origin, platform and other systems and decimal numbers as part of a rating system. From there we can create a model that has a binary of liked vs hated with a cutoff of 5.0. This model can then predict the most likely game a avid gamer may choose based on their likes and dislikes. This could be helpful as part of a recommendation system on online gaming platforms and tech websites.