---
title: 'Word2Vec'
author: "STUDENT NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries + functions

Load all the libraries or functions that you will use to for the rest of the assignment. It is helpful to define your libraries and functions at the top of a report, so that others can know what they need for the report to compile correctly.

```{r libaries}
##r chunk
```

Load the Python libraries or functions that you will use for that section. 

```{python}
##python chunk

```

## The Data

The dataset is a set of Youtube comments that have been coded as:
  - 1: spam youtube messages 
  - 0: good youtube messages
  - This data is stored in the `CLASS` column

Import the data using either R or Python. I put a Python chunk here because you will need one to import the data, but if you want to first import into R, that's fine. 

```{python}
##python chunk

```

## Clean up the data

Use one of our clean text functions to clean up the `CONTENT` column in the dataset. 

```{python}
##python chunk

```

## Split the data

Split the data into testing and training data.

```{python}
##python chunk

```

## Process the data

For FastText OR word2vec, create the tokenized vectors of the text.

```{python}
##python chunk

```

## Build Model

Build either a word2vec or FastText model.

```{python}
##python chunk

```

## Convert the model

Convert the model into a set of features to use in our classifier. 

```{python}
##python chunk

```

## Build a classifier model

In class, we used logistic regression to classify the data. You can use any machine learning algorithm you want here, and build a classification model. 

```{python}
##python chunk

```

## Examine the results

Print out the accuracy, recall, and precision of your model.

```{python}
##python chunk

```

## Discussion Questions

- How well did the model you build do? What could you do to improve the accuracy of the model? (Interpret the model in at least 5 sentences)

  - ANSWER: 

- Describe a set of texts and research question that interests you that could be explored using this method. Basically, what is a potential application of this method to another area of research?

  - ANSWER: 