---
title             : "Internet Culture – Incel Language Detection within in-groups on the internet"
shorttitle        : "Incel Language Detection"

author: 
  - name          : "Akane Simpson"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "326 Market St, Harrisburg, PA 17101"
    email         : "asimpson@my.harrisburgu.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Harrisburg University of Science & Technology"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}
    
    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

This is the initial line for an introduction to the paper. In this paper I will seek to analyze and document an new discovery about our current society.

# Literature Review
This is the first line for a literature review.

@helm2022examining opening line says, "The online presence of incels, or involuntary celibates, has been an increasing security concern for researchers, practitioners, and policymakers in recent years..." while @o2022exploration suports this claim "...several violent incidents have been linked to this community, which suggests that incel communities may resemble other ideologically motivated extremist groups".

# Methods
In this section I will talk about how I planned my study by calculating the required sample size, the data that I was able to find or generate, list and discuss how this relates to answering our hypothesis or whatever.

## Participants
This study should have no participants needed.

## Procedure
The study will use a step by step procedure to identify, measure and observe interactions by users on the internet.

## Measures
The measures that we identify are language and sentiment in these online posts.

## Analysis
We used `r cite_r("r-references.bib")` for all our analyses.
In our analysis we will see if there is a established connection that helps prove or disprove our prior hypothesis.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
