---
title: "Module 2 - Working with Categorical Data 1"
author: "Akane Simpson"
date: '2022-03-30'
output: html_document
---

```{r setup, include=FALSE}
setwd("~/hu_lectures/anly_545/Lecture 2/")
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

```{r Basic Data Wrangling, include=FALSE}
base_form <- read.csv('./data.2.csv')

summary(base_form)
str(base_form)

total_cases <- nrow(base_form)

cleaned <- base_form[,-6]
cleaned$stars <- factor(cleaned$stars)
cleaned$race <- factor(cleaned$race)
cleaned$restaurant <- factor(cleaned$restaurant)
cleaned$sex <- factor(cleaned$sex)

summary(cleaned)

restrnt.names <- levels(cleaned$restaurant)
restrnt.count <- cleaned %>% 
  group_by(restaurant) %>% 
  tally()
restrnt.count <- restrnt.count$n
restrnt.percnt <-  round((restrnt.count/total_cases)*100, 1)

rest.freq <- data.frame(restaurant = restrnt.names,
              count = restrnt.count,
              percentage = restrnt.percnt)
rest.freq <- rest.freq %>% 
  arrange(., desc(count))
```

## Analysis (Opinion)

1. The restaurants are displayed in the order they are because <u>they were ordered
by the highest count (descending order). This allows readers to readily identify
restaurants with the most customers</u>.

2. The addition of the percentage column to the table allows the reader to see 
<u>the distribution of customers between each restaurant</u>.

## Table and Figures

Find the total number of cases: <u>`r total_cases`</u>

For each restaurant, find the total number of customers.

```{r Restaurant Table, include=FALSE}
table <- knitr::kable(rest.freq, caption = "Restaurant Customers", col.names = c("Restaurant", "Count", "Percentage"))
```
`r table`




