---
title: "Module 2 - Working with Categorical Data 2"
author: "Akane Simpson"
date: '2022-04-02'
output: html_document
---

```{r setup, include=FALSE}
setwd("~/hu_lectures/anly_545/Lecture 2/")
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(kableExtra)
library(reshape2)

options(knitr.table.format = "html")
```

## Analysis (Opinion)

In the 2nd and 3rd tables:

1. The restaurants are displayed in the order they are because <u>the are arranged by the largest value in the overall column</u>.
2. The addition of the overall column to the table allows the reader to see <u>the overall column shows us the total share of participants tend to eat at a specific restaurant</u>.
3. The position of the overall column is important because <u>it will give us a better idea which restaurant has a larger customer base from a glance.</u>.
4. The groupings of the sex and race columns are important because <u>it helps separate the overall value into a more detailed breakdown of sample representation with race being used to show which races or sexes tend to eat more at a specific restaurant</u>.


```{r Basic Data Wrangling, include=FALSE}
base <- read.csv('./data.2.csv')

summary(base)
str(base)


cleaned <- base[,-6]
cleaned$race <- factor(cleaned$race)
cleaned$restaurant <- factor(cleaned$restaurant)
cleaned$sex <- factor(cleaned$sex)

total_cases <- nrow(cleaned)
rest_sex_table <- table(cleaned$sex)
total_male <-  as.data.frame(rest_sex_table)[2,2]
total_female <- as.data.frame(rest_sex_table)[1,2]

restrnt.names <- levels(cleaned$restaurant)
restrnt.sex <- levels(cleaned$sex)
restrnt.race <- levels(cleaned$race)
```

## Table and Figures
Find the total number of cases: <u>`r total_cases`</u>

For each restaurant, find the total number of customers.

```{r Restaurant Count, include=FALSE}

restrnt.count <- cleaned %>% 
  group_by(restaurant) %>% 
  tally()

restrnt.count <- restrnt.count$n
restrnt.percnt <-  round((restrnt.count/total_cases)*100, 2)

rest.freq <- data.frame(restaurant = restrnt.names,
              count = restrnt.count,
              percentage = restrnt.percnt)

rest.freq <- rest.freq %>% 
  arrange(., desc(count))

table <- kbl(rest.freq, caption = "Restaurant Customers", col.names = c("Restaurant", "Count", "Percentage")) %>% 
  kable_styling()
```

`r table`

For each restaurant, find the average star rating by __Sex__.

```{r Average Rating by Sex, include=FALSE}
restrnt.ratings <- cleaned %>% 
  group_by(restaurant, sex) %>% 
  summarise(avg = round(mean(stars), 2)) 

restrnt.overall <- restrnt.ratings %>% 
  summarise(overall = round(sum(avg)/length(restrnt.sex), 2)) 

restrnt.ratings <- dcast(restrnt.ratings, restaurant ~ sex, value.var = "avg")

restrnt.freq <- inner_join(restrnt.overall, restrnt.ratings, by = "restaurant") %>% 
  arrange(., desc(overall))

table <- kbl(restrnt.freq, caption = "Average Customer Rating by Sex", 
                      col.names = c("Restaurant", "Overall", restrnt.sex)) %>% 
  add_header_above(c(" " = 2, "Sex" = length(restrnt.sex))) %>% 
  kable_styling()
```

`r table`

For each restaurant, find the average star rating by __Race__.

```{r Average Rating by Race, include=FALSE}
restrnt.ratings <- cleaned %>% 
  group_by(restaurant, race) %>% 
  summarise(avg = round(mean(stars), 2)) 

restrnt.overall <- restrnt.ratings %>% 
  summarise(overall = round(sum(avg)/length(restrnt.race), 2))

restrnt.ratings <- dcast(restrnt.ratings, restaurant ~ race, value.var = "avg")

restrnt.freq <- inner_join(restrnt.overall, restrnt.ratings, by = "restaurant") %>% 
  arrange(., desc(overall))

table <- kbl(restrnt.freq, caption = "Average Customer Rating by Sex", 
                      col.names = c("Restaurant", "Overall", restrnt.race)) %>% 
  add_header_above(c(" " = 2, "Race" = length(restrnt.race))) %>% 
  kable_styling()
```

`r table`
