---
title: "Data Cleaning and Clustering"
author: "Yuan Ji"
date: "2022/1/28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(readxl)
library(stringr)
library(dplyr)  
library(dendextend)
library(data.table) 
library(ggplot2)
```

## Load Raw COVID Data (With Data Cleaning)

Load Raw COVID-19 cases data 
```{r load data}
covid_data <- read.csv("covid_data.csv")
```

Remove data where continent is empty, save it as COVID time series data for later use
```{r data cleaning}
covid_data_ts <- covid_data[covid_data$continent!='',]
```

## Clustering Analysis

Generate country information summary data set, which will be used for clustering analysis
```{r summary}
country_col <- c( 'continent',
                  'location',
                  'population',
                  'population_density',
                  'median_age',
                  'aged_65_older',
                  'aged_70_older',
                  'gdp_per_capita',
                  'extreme_poverty',
                  'cardiovasc_death_rate',
                  'diabetes_prevalence',
                  'female_smokers',
                  'male_smokers',
                  'handwashing_facilities',
                  'hospital_beds_per_thousand',
                  'life_expectancy',
                  'human_development_index')

covid_data_country <- covid_data_ts [,country_col]
covid_data_country <- distinct(covid_data_country)

#summary(covid_data_country)

# hand washing facilities and extreme property are the attributes which have too many NAs, we remove this features
country_col <- c( 'continent',
                  'location',
                  'population',
                  'population_density',
                  'median_age',
                  'aged_65_older',
                  'aged_70_older',
                  'gdp_per_capita',
                  'cardiovasc_death_rate',
                  'diabetes_prevalence',
                  'female_smokers',
                  'male_smokers',
                  'hospital_beds_per_thousand',
                  'life_expectancy',
                  'human_development_index')
covid_data_country <- covid_data_country[,country_col]

# Exam the percentages for missing values for each country
missSum <- function(x){ sum(is.na(x)) }
missing <- apply(covid_data_country, 1, missSum)

# Final Country data set with no missing values
covid_data_country <- subset(covid_data_country, missing <=0)

summary(covid_data_country)
```
Since we have over one hundred countries remain in our dataset after data cleaning, we have to further explore the relationship between countries by clustering them into groups. Then it would be easier to pick a few pairs of countries to compare their daily active COVID cases, vaccine rate, etc.

K-Means Clustering whole procedure
```{r}
set.seed(1111)

feature_col <- c( 'population',
                  'population_density',
                  'median_age',
                  'aged_65_older',
                  'aged_70_older',
                  'gdp_per_capita',
                  'cardiovasc_death_rate',
                  'diabetes_prevalence',
                  'female_smokers',
                  'male_smokers',
                  'hospital_beds_per_thousand',
                  'life_expectancy',
                  'human_development_index')

country_kmean = covid_data_country[,-c(1,2)]


#standardize the variables
country_standard <- country_kmean %>% mutate_at(feature_col, ~(scale(.) %>% as.vector))

# K mean algorithm applying
kmeansObj <- kmeans(country_standard,iter.max = 500, centers=4)
# Add the Kmean group result into country data
covid_data_country$group = kmeansObj$cluster

# Check how many countries are there in each group
kmeansObj$size

#group by continent and country

country_kmean_group <-covid_data_country %>%
                     group_by(group) %>%
                     summarise_at(vars(population,
                                       population_density,
                                       median_age,
                                       aged_65_older,
                                       aged_70_older,
                                       gdp_per_capita,
                                       cardiovasc_death_rate,
                                       diabetes_prevalence,
                                       female_smokers,
                                       male_smokers,
                                       hospital_beds_per_thousand,
                                       life_expectancy,
                                       human_development_index), 
                     list(name = mean))


```


Pick the 5 countries with highest population from each group to get my finalized country list (Result get from the data set Covid_data_country)
Group1 : United States, Japan, Germany, United Kingdom, France
Group2 : China, India, Indonesia, Bangladesh, Russia
Group3 : Pakistan, Ethiopia, Tanzania ,  South Africa,Kenya,
Group4 : Brazil, Mexico, Egypt, Turkey, Iran

```{r filtering country}
country_list <- c('United States', 
                  'Japan', 
                  'Germany', 
                  'United Kingdom', 
                  'France',
                  'China', 
                  'India', 
                  'Indonesia', 
                  'Bangladesh', 
                  'Russia',
                  'Pakistan', 
                  'Ethiopia', 
                  'Tanzania', 
                  'Kenya',
                  'South Africa',
                  'Brazil',
                  'Mexico', 
                  'Egypt', 
                  'Turkey', 
                  'Iran')

covid_data_country_select <-covid_data_country[
                         covid_data_country$location %in% country_list,] 
  #                       

```



Hierarchial Clutering analysis for countries to further analysis the result


```{r clustering}

#Pre-process data
#Use person names as row names - creates plot labels
row.names(covid_data_country_select) = covid_data_country_select$location

#Drop first two columns
covid_data_country_hclust = covid_data_country_select[,-c(1,2)]
#standardize the variables
country_standard_hclust <- covid_data_country_hclust %>% mutate_at(feature_col, ~(scale(.) %>% as.vector))

#Hierarchical Clustering

#Calculate dendrogram
dendrogram =as.dendrogram(
      hclust(
        dist(
        country_standard_hclust
      )
    )
  )
#Plot dendrogram
plot(dendrogram,main="Dendrogram for clustering countries",ylab= 'Distances')

```
## Time Series Plotting

Some initial analysis by plotting Time Series Analysis between pairs of countries based on clustering result


Pair One : UK and France (compare within same groups)
```{r uk vs france}
#Plot new cases
sub <- covid_data_ts[covid_data_ts$location %in% c('United Kingdom','France')
                     ,c('date','location','new_cases_smoothed','new_vaccinations_smoothed')]
sub$date <- as.Date(sub$date, format="%Y-%m-%d")

sub <- sub[(sub$date>"2020-03-01") &(sub$date<="2021-12-31"),]

# Plot for daily new cases

new_cases_plot <- ggplot(sub, aes(x=date, y=new_cases_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'Active New Cases', title ='Smoothed Daily New Cases (7 days average)')
  
new_cases_plot

# Plot for daily Vaccinations
sub <- sub[(sub$date>"2021-01-01") &(sub$date<="2021-12-31"),]
new_vacc_plot <- ggplot(sub, aes(x=date, y=new_vaccinations_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'New Vaccinations Amount', title ='Smoothed Daily New Vaccinations (7 days average)')
  
new_vacc_plot
```


Pair Two : USA and Indonesia (compare between two countries with similar population but in different group)

```{r usa vs Indonesia}
#Plot new cases
sub <- covid_data_ts[covid_data_ts$location %in% c('United States','Indonesia')
                     ,c('date','location','new_cases_smoothed','new_vaccinations_smoothed')]
sub$date <- as.Date(sub$date, format="%Y-%m-%d")

sub <- sub[(sub$date>"2020-03-01") &(sub$date<="2021-12-31"),]

# Plot for daily new cases

new_cases_plot <- ggplot(sub, aes(x=date, y=new_cases_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'Active New Cases', title ='Smoothed Daily New Cases (7 days average)')
  
new_cases_plot

# Plot for daily Vaccinations
sub <- sub[(sub$date>"2021-01-01") &(sub$date<="2021-12-31"),]
new_vacc_plot <- ggplot(sub, aes(x=date, y=new_vaccinations_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'New Vaccinations Amount', title ='Smoothed Daily New Vaccinations (7 days average)')
  
new_vacc_plot

```

Pair Three : Japan and Iran ( compare two countries with the max distance from dendrogram)

```{r japan vs Iran}
#Plot new cases
sub <- covid_data_ts[covid_data_ts$location %in% c('Japan','Iran')
                     ,c('date','location','new_cases_smoothed','new_vaccinations_smoothed')]
sub$date <- as.Date(sub$date, format="%Y-%m-%d")

sub <- sub[(sub$date>"2020-03-01") &(sub$date<="2021-12-31"),]

# Plot for daily new cases

new_cases_plot <- ggplot(sub, aes(x=date, y=new_cases_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'Active New Cases', title ='Smoothed Daily New Cases (7 days average)')
  
new_cases_plot

# Plot for daily Vaccinations
sub <- sub[(sub$date>"2021-01-01") &(sub$date<="2021-12-31"),]
new_vacc_plot <- ggplot(sub, aes(x=date, y=new_vaccinations_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'New Vaccinations Amount', title ='Smoothed Daily New Vaccinations (7 days average)')
  
new_vacc_plot

```
```{r}


```
There could be other pairs/groups of countries we are interested to research for the next step.
