---
title: "Data Cleaning, EDA and Clustering"
author: "Yuan Ji, Yingzhi Wang"
date: "2022/1/30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(readxl)
library(stringr)
library(dplyr)  
library(dendextend)
library(data.table) 
library(ggplot2)
library(tidyr)
library(plotly)
```

## Load Raw COVID Data (With Data Cleaning)

Load Raw COVID-19 cases data 
```{r load data}
setwd('~/hu_lectures/anly_506/Project')
covid_data <- read.csv("covid_data.csv")
```

Remove data where continent is empty, save it as COVID time series data for later use
```{r data cleaning}
covid_data_ts <- covid_data[covid_data$continent!='',] # country-level data
covid_data_cn <- covid_data[covid_data$continent=='',] # worlf-level data
```
## Exploratory Data Analysis
```{r}
covid_data_cn$date <- as.Date(covid_data_cn$date, format="%Y-%m-%d")
covid_data_ts$date <- as.Date(covid_data_ts$date, format="%Y-%m-%d")
covid_data_cn <- covid_data_cn[(covid_data_cn$date >= "2020-03-01") &(covid_data_cn$date<="2021-12-31"),] # use the data from 3/1/2020 to 12/31/2021
covid_data_ts <- covid_data_ts[(covid_data_ts$date >= "2020-03-01") &(covid_data_ts$date<="2021-12-31"),] # use the data from 3/1/2020 to 12/31/2021
```

Under the world-level data, we can look at the trend of total cases group by different continents:

```{r}
# Plot for total cases btw different continents
regions <- covid_data_cn[covid_data_cn$location %in% c('Africa','Asia', 'Europe', 'North America', 'South America', 'Oceania', 'World'),]
total_cases_plot <- ggplot(regions, aes(x=date, y=total_cases, color=location)) +
  geom_line() + 
  theme_bw() +
  labs(x = 'Date', y = 'Total Cases', title ='Total Cases per Continent')
total_cases_plot
```


```{r}
# World total cases/death
world_data <- covid_data_cn[covid_data_cn$location %in% c('World'),]

world_total_cases <- world_data %>%
  ggplot(aes(x = date, y = total_cases))+
  geom_line(color = 'orange') +
  theme(legend.position = "none") +
  geom_line(aes(x = date, y = total_deaths * 100), color = "red") +
  scale_y_continuous(
    labels = scales::comma,
    name = "Cases",
    sec.axis = sec_axis(total_deaths ~ . / 100,
                        name = "Deaths",
                        labels = scales::comma
    )
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_text(color = "orange", size = 13),
    axis.title.y.right = element_text(color = "red", size = 13)
  ) +
  labs(
    title = "World Total Cases vs. Deaths",
    x = "Date"
  )
show(world_total_cases)

```

If we take a look at the new cases/deaths per day, we can notice: 
1. When cases start to rise, deaths follow with a lag. 
2. We have had three spikes in cases so far and in each successive instance the mortality has risen by a smaller amount. This suggests that, thankfully, we are getting better at treating this disease. It is NOT a function of increased testing because positivity rates have not been falling.
```{r}
# World New Cases/death
world_new_cases <- world_data %>%
  ggplot(aes(x = date, y = new_cases_smoothed))+
  geom_line(color = 'orange') +
  theme(legend.position = "none") +
  geom_line(aes(x = date, y = new_deaths_smoothed * 50), color = "red") +
  scale_y_continuous(
    labels = scales::comma,
    name = "Cases",
    sec.axis = sec_axis(new_deaths_smoothed ~ . / 50,
                        name = "Deaths",
                        labels = scales::comma
    )
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_text(color = "orange", size = 13),
    axis.title.y.right = element_text(color = "red", size = 13)
  ) +
  labs(
    title = "World New Cases vs. Deaths",
    x = "Date"
  )
world_new_cases
```

We can also take a look at the world mortality rate since the pandemic started.
1. There was a sudden drop in the reproduction rate at the start of the pandemic
2. The mortality rate of the pandemic is steadily rising
3. The birth rate is slowly rising back to pre-pandemic levels
```{r}
# World Mortality/Reproduction Rate per 100,000
world_mortality_rate <- world_data %>%
  ggplot(aes(x = date, y = (total_deaths_per_million / population) * 10^5))+
  geom_line(color = 'firebrick') +
  theme(legend.position = "none") +
  geom_line(aes(x = date, y = reproduction_rate), color = "cadetblue") +
  scale_y_continuous(
    labels = scales::comma,
    name = "Reproduction Rate",
    sec.axis =  sec_axis( reproduction_rate ~ .,
                         name = "Mortality Rate",
                         labels = scales::comma
               )
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_text(color = "cadetblue", size = 13),
    axis.title.y.right = element_text(color = "firebrick", size = 13)
  ) +
  labs(
    title = "World Mortality Rate vs Reproduction Rate per 100,000",
    x = "Date"
  )
world_mortality_rate
```


```{r}
# World Mortality/Case Mortality
world_case_and_mortality_rate <- world_data %>%
  ggplot(aes(x = date, y = (total_cases_per_million / population) * 10^5))+
  geom_line(color = 'darkgoldenrod') +
  theme(legend.position = "none") +
  geom_line(aes(x = date, y = (total_deaths_per_million / population) * 10^5), color = "firebrick") +
  scale_y_continuous(
    labels = scales::comma,
    name = "Case Rate",
    sec.axis =  sec_axis( (total_deaths / population) ~ . ,
                         name = "Mortality Rate",
                         labels = scales::comma
               )
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_text(color = "darkgoldenrod", size = 13),
    axis.title.y.right = element_text(color = "firebrick", size = 13)
  ) +
  labs(
    title = "World Case Rates vs Mortality Rate per 100,000",
    x = "Date"
  )
world_case_and_mortality_rate
```


Under Country level, let's take a look at the summary of top 20 total confirmed cases group by countries:
```{r}
summary_ts <- covid_data_ts %>%
  mutate(Country = location) %>% 
  group_by(Country) %>%
  summarise(Total_cases_Sum = sum(new_cases)) %>%
  arrange(-Total_cases_Sum)
summary_ts%>% head(20)
```

Plot the top 5 countries with confirmed cases:
```{r}
top_5_country <- c('United States', 
                  'India', 
                  'United Kingdom', 
                  'Russia',
                  'Germany')
top_5_data <-covid_data_ts[
                         covid_data_ts$location %in% top_5_country,] 
top_5_data %>% 
  mutate(Country = location) %>% 
  plot_ly(x = ~date, y = ~total_cases, color = ~Country, type = 'scatter', mode = 'lines') %>% 
  layout(
    yaxis = list(title = "Confirmed Cases"),
    xaxis = list(title = "Date"),
    title = "Top 5 Countries with Covid19 Confirmed Cases",
    legend = list(x = 100, y = 0.5),
    hovermode = "compare"
  )
```

Summary of new cases/deaths at the last day by country (as of 2021-12-31):

```{r}
covid_data_ts %>%
  filter(date == max(date)) %>% 
  mutate(Country = location) %>% 
  select(Country, new_cases, new_deaths) %>% 
  arrange(-new_cases)
```

What's the contribution of each country on the total cases in the world?

```{r}
tree_country_sum <- covid_data_ts %>%
  group_by(location) %>%
  summarise(Total_cases_Sum = sum(new_cases)) %>%
  arrange(-Total_cases_Sum) %>% 
  mutate(parents = 'Total Cases') %>% 
  ungroup() 

plot_ly(data = tree_country_sum,
        type= "treemap",
        values = ~Total_cases_Sum,
        labels= ~ location,
        parents=  ~parents,
        domain = list(column=0),
        name = "Total Cases",
        textinfo="label+value+percent parent")
```

Since we have vaccination during 2021, what is the top 20 countries which have the fully vaccinated rate per population?
```{r}
# the top 20 countries which people fully vaccinated
vaccine <- covid_data_ts %>% 
  filter(date == max(date),
         !is.na(population)) %>% 
  mutate(fully_vaccinated_ratio = people_fully_vaccinated / population) %>%
  arrange(- fully_vaccinated_ratio) %>%
  head(20) %>%
  arrange(fully_vaccinated_ratio) %>%
  mutate(country = factor(location, levels = location)) %>%
  plot_ly(y = ~ country,
          x = ~ round(100 * fully_vaccinated_ratio, 2),
          text = ~ paste(round(100 * fully_vaccinated_ratio, 1), "%"),
          textposition = 'auto',
          orientation = "h",
          type = "bar") %>%
  layout(title = "Percentage of Fully Vaccineted Population - Top 20 Countries",
         yaxis = list(title = ""),
         xaxis = list(title = "Fully Vaccinated Ratio",
                      ticksuffix = "%"))
vaccine

```


## Clustering Analysis

Generate country information summary data set, which will be used for clustering analysis
```{r summary}
country_col <- c( 'continent',
                  'location',
                  'population',
                  'population_density',
                  'median_age',
                  'aged_65_older',
                  'aged_70_older',
                  'gdp_per_capita',
                  'extreme_poverty',
                  'cardiovasc_death_rate',
                  'diabetes_prevalence',
                  'female_smokers',
                  'male_smokers',
                  'handwashing_facilities',
                  'hospital_beds_per_thousand',
                  'life_expectancy',
                  'human_development_index')

covid_data_country <- covid_data_ts [,country_col]
covid_data_country <- distinct(covid_data_country)

#summary(covid_data_country)

# hand washing facilities and extreme property are the attributes which have too many NAs, we remove this features
country_col <- c( 'continent',
                  'location',
                  'population',
                  'population_density',
                  'median_age',
                  'aged_65_older',
                  'aged_70_older',
                  'gdp_per_capita',
                  'cardiovasc_death_rate',
                  'diabetes_prevalence',
                  'female_smokers',
                  'male_smokers',
                  'hospital_beds_per_thousand',
                  'life_expectancy',
                  'human_development_index')
covid_data_country <- covid_data_country[,country_col]

# Exam the percentages for missing values for each country
missSum <- function(x){ sum(is.na(x)) }
missing <- apply(covid_data_country, 1, missSum)

# Final Country data set with no missing values
covid_data_country <- subset(covid_data_country, missing <=0)

summary(covid_data_country)
```
Since we have over one hundred countries remain in our dataset after data cleaning, we have to further explore the relationship between countries by clustering them into groups. Then it would be easier to pick a few pairs of countries to compare their daily active COVID cases, vaccine rate, etc.

K-Means Clustering whole procedure
```{r}
set.seed(1111)

feature_col <- c( 'population',
                  'population_density',
                  'median_age',
                  'aged_65_older',
                  'aged_70_older',
                  'gdp_per_capita',
                  'cardiovasc_death_rate',
                  'diabetes_prevalence',
                  'female_smokers',
                  'male_smokers',
                  'hospital_beds_per_thousand',
                  'life_expectancy',
                  'human_development_index')

country_kmean = covid_data_country[,-c(1,2)]


#standardize the variables
country_standard <- country_kmean %>% mutate_at(feature_col, ~(scale(.) %>% as.vector))

# K mean algorithm applying
kmeansObj <- kmeans(country_standard,iter.max = 500, centers=4)
# Add the Kmean group result into country data
covid_data_country$group = kmeansObj$cluster

# Check how many countries are there in each group
kmeansObj$size

#group by continent and country

country_kmean_group <-covid_data_country %>%
                     group_by(group) %>%
                     summarise_at(vars(population,
                                       population_density,
                                       median_age,
                                       aged_65_older,
                                       aged_70_older,
                                       gdp_per_capita,
                                       cardiovasc_death_rate,
                                       diabetes_prevalence,
                                       female_smokers,
                                       male_smokers,
                                       hospital_beds_per_thousand,
                                       life_expectancy,
                                       human_development_index), 
                     list(name = mean))


```


Pick the 5 countries with highest population from each group to get my finalized country list (Result get from the data set Covid_data_country)
Group1 : United States, Japan, Germany, United Kingdom, France
Group2 : China, India, Indonesia, Bangladesh, Russia
Group3 : Pakistan, Ethiopia, Tanzania ,  South Africa,Kenya,
Group4 : Brazil, Mexico, Egypt, Turkey, Iran

```{r filtering country}
country_list <- c('United States', 
                  'Japan', 
                  'Germany', 
                  'United Kingdom', 
                  'France',
                  'China', 
                  'India', 
                  'Indonesia', 
                  'Bangladesh', 
                  'Russia',
                  'Pakistan', 
                  'Ethiopia', 
                  'Tanzania', 
                  'Kenya',
                  'South Africa',
                  'Brazil',
                  'Mexico', 
                  'Egypt', 
                  'Turkey', 
                  'Iran')

covid_data_country_select <-covid_data_country[
                         covid_data_country$location %in% country_list,] 
  #                       

```



Hierarchial Clutering analysis for countries to further analysis the result


```{r clustering}

#Pre-process data
#Use person names as row names - creates plot labels
row.names(covid_data_country_select) = covid_data_country_select$location

#Drop first two columns
covid_data_country_hclust = covid_data_country_select[,-c(1,2)]
#standardize the variables
country_standard_hclust <- covid_data_country_hclust %>% mutate_at(feature_col, ~(scale(.) %>% as.vector))

#Hierarchical Clustering

#Calculate dendrogram
dendrogram =as.dendrogram(
      hclust(
        dist(
        country_standard_hclust
      )
    )
  )
#Plot dendrogram
plot(dendrogram,main="Dendrogram for clustering countries",ylab= 'Distances')

```
## Time Series Plotting

Some initial analysis by plotting Time Series Analysis between pairs of countries based on clustering result


Pair One : UK and France (compare within same groups)
```{r uk vs france}
#Plot new cases
sub <- covid_data_ts[covid_data_ts$location %in% c('United Kingdom','France')
                     ,c('date','location','new_cases_smoothed','new_vaccinations_smoothed')]
sub$date <- as.Date(sub$date, format="%Y-%m-%d")

sub <- sub[(sub$date>"2020-03-01") &(sub$date<="2021-12-31"),]

# Plot for daily new cases

new_cases_plot <- ggplot(sub, aes(x=date, y=new_cases_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'Active New Cases', title ='Smoothed Daily New Cases (7 days average)')
  
new_cases_plot

# Plot for daily Vaccinations
sub <- sub[(sub$date>"2021-01-01") &(sub$date<="2021-12-31"),]
new_vacc_plot <- ggplot(sub, aes(x=date, y=new_vaccinations_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'New Vaccinations Amount', title ='Smoothed Daily New Vaccinations (7 days average)')
  
new_vacc_plot
```


Pair Two : USA and Indonesia (compare between two countries with similar population but in different group)

```{r usa vs Indonesia}
#Plot new cases
sub <- covid_data_ts[covid_data_ts$location %in% c('United States','Indonesia')
                     ,c('date','location','new_cases_smoothed','new_vaccinations_smoothed')]
sub$date <- as.Date(sub$date, format="%Y-%m-%d")

sub <- sub[(sub$date>"2020-03-01") &(sub$date<="2021-12-31"),]

# Plot for daily new cases

new_cases_plot <- ggplot(sub, aes(x=date, y=new_cases_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'Active New Cases', title ='Smoothed Daily New Cases (7 days average)')
  
new_cases_plot

# Plot for daily Vaccinations
sub <- sub[(sub$date>"2021-01-01") &(sub$date<="2021-12-31"),]
new_vacc_plot <- ggplot(sub, aes(x=date, y=new_vaccinations_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'New Vaccinations Amount', title ='Smoothed Daily New Vaccinations (7 days average)')
  
new_vacc_plot

```

Pair Three : Japan and Iran ( compare two countries with the max distance from dendrogram)

```{r japan vs Iran}
#Plot new cases
sub <- covid_data_ts[covid_data_ts$location %in% c('Japan','Iran')
                     ,c('date','location','new_cases_smoothed','new_vaccinations_smoothed')]
sub$date <- as.Date(sub$date, format="%Y-%m-%d")

sub <- sub[(sub$date>"2020-03-01") &(sub$date<="2021-12-31"),]

# Plot for daily new cases

new_cases_plot <- ggplot(sub, aes(x=date, y=new_cases_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'Active New Cases', title ='Smoothed Daily New Cases (7 days average)')
  
new_cases_plot

# Plot for daily Vaccinations
sub <- sub[(sub$date>"2021-01-01") &(sub$date<="2021-12-31"),]
new_vacc_plot <- ggplot(sub, aes(x=date, y=new_vaccinations_smoothed, color=location)) +
  geom_line() + 
  labs(x = 'Date', y = 'New Vaccinations Amount', title ='Smoothed Daily New Vaccinations (7 days average)')
  
new_vacc_plot

```
```{r}
```
There could be other pairs/groups of countries we are interested to research for the next step.
