---
title: "Lecture 3: Assignment 3 - Probability"
author: "Akane Simpson"
date: '`r Sys.Date()`'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load('./kobe.RData')
```

## NOW ON YOUR OWN

```{r Kobe Basketball Streak}
  kobe_streak <- calc_streak(kobe$basket)
  table(kobe_streak)/100
  
  barplot(table(kobe_streak), xlab = "Hot Streak", ylab = "Frequency", ylim = c(0,50), main = "Kob Bryant's Hot Streak Calculation (2009)", col = "gold")
    
  kobe_shot_rate <- table(kobe$basket)/133
```


```{r Creating a Simulated Basket Sample}
  outcomes <- c("H", "M")
  sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(.5, 0.5))
```


1. Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? 
```{r}
sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(0.45,.55))
sim_streak <- calc_streak(sim_basket)

sim_basket_shot_rate <- table(sim_basket)/133
table(calc_streak(sim_basket))/100

barplot(table(sim_streak), xlab = "Hot Streak", ylab = 'Frequency', ylim = c(0,50),main = " Simulated Shooter's Hot Streak", col = "purple")
```

```{r,echo=FALSE}
print("The simulated basket rate streak is right tailed and has a very high kurtosis. It is similar to the streak of Kobe with a high of a '0' streak and as the streak grows the rate diminishes.")
```

  a. How long is the player's longest streak of baskets in 133 shots?
```{r,echo=FALSE}
sprintf("The longest streak the simulated player had was at a streak of '0' with %d individual shots. The second longest streak for the player is '1' with a total of %d individual shots on target.", table(sim_streak)["0"], table(sim_streak)["1"])
```
       
       
2. If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.
```{r, echo=FALSE}
print("If the simulation was ran again then the results would be somewhat similar. AN explanation for this change is how the probability of H and M are assinged. With each independent shot the simualted player is more likely to miss but the rate of streaks can increase. This streak amount could increase to over 4 while the probability of this happening will diminish.")
```

       
3. How does Kobe Bryant's distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe's shooting patterns? Explain.
```{r,echo=FALSE}
sprintf("Kobe's streak and the current sample share the distribution shape where boith are right skewed with a high kurtosis. However, the data presented does not support a hot hand model as both players are more likely to miss their shots and are less likely to continue a streak over 2. To reinforce this point both Kobe and the simulated player have a hit rate of %.2f%% and %.2f%% respectively. This means each individual shot has an almost 50/50 chance of hitting the basket for both players.", kobe_shot_rate["H"], sim_basket_shot_rate["H"])
```
