---
title: 'Assignment #1'
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Chapter 2 - Large Worlds and Small Worlds

The objectives of this problem set is to work with the conceptual mechanics of Bayesian data analysis.  The target of inference in Bayesian inference is a posterior probability distribution. Posterior probabilities state the relative numbers of ways each conjectured cause of the data could have produced the data. These relative numbers indicate plausibilities of the different conjectures. These plausibilities are updated in light of observations through Bayesian updating.  

Place each answer inside the code chunk (grey box).  The code chunks should contain a text response or a code that completes/answers the question or activity requested. Make sure to include plots if the question requests them.  

Finally, upon completion, name your final output `.html` file as: `YourName_ANLY505-Year-Semester.html` and publish the assignment to your R Pubs account and submit the link to Canvas. Each question is worth 5 points

## Questions

**2-1.** Suppose you have a deck with only three cards. Each card has two sides, and each side is either
black or white. One card has two black sides. The second card has one black and one white side. The
third card has two white sides. Now suppose all three cards are placed in a bag and shuffled. Someone
reaches into the bag and pulls out a card and places it flat on a table. A black side is shown facing up,
but you don’t know the color of the side facing down. Show that the probability that the other side is
also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This
means counting up the ways that each card could produce the observed data (a black side facing up
on the table).

```{r 2-1}

```

**2-2.** Now suppose there are four cards: B/B, B/W, W/W, and another B/B. Again suppose a card is
drawn from the bag and a black side appears face up. Again calculate the probability that the other
side is black.

```{r 2-2}

```

**2-3.** Imagine that black ink is heavy, and so cards with black sides are heavier than cards with white
sides. As a result, it’s less likely that a card with black sides is pulled from the bag. So again assume
there are three cards: B/B, B/W, and W/W. After experimenting a number of times, you conclude that
for every way to pull the B/B card from the bag, there are 2 ways to pull the B/W card and 3 ways to
pull the W/W card. Again suppose that a card is pulled and a black side appears face up. Show that
the probability the other side is black is now 0.5. Use the counting method, as before.

```{r 2-3}

```

**2-4.** Assume again the original card problem, with a single card showing a black side face up. Before
looking at the other side, we draw another card from the bag and lay it face up on the table. The face
that is shown on the new card is white. Show that the probability that the first card, the one showing
a black side, has black on its other side is now 0.75. Use the counting method, if you can. Hint: Treat
this like the sequence of globe tosses, counting all the ways to see each observation, for each possible
first card.

```{r 2-4}

```

**2-5.** Suppose there are two species of panda bear. Both are equally common in the wild and live
in the same places. They look exactly alike and eat the same food, and there is yet no genetic assay
capable of telling them apart. They differ however in their family sizes. Species A gives birth to twins
10% of the time, otherwise birthing a single infant. Species B births twins 20% of the time, otherwise
birthing singleton infants. Assume these numbers are known with certainty, from many years of field
research.
Now suppose you are managing a captive panda breeding program. You have a new female panda
of unknown species, and she has just given birth to twins. What is the probability that her next birth
will also be twins?

```{r 2-5}

```

